---
title: "BC Crime Data - Exploration Worksheet"
author: "Casper Kacper Ruta"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: flatly
    code_folding: show
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 9,
  fig.height = 5
)
```

# Purpose

This worksheet is a companion to the main analysis report (`BC-CRIME.Rmd`). It
provides a quick-start template for ad-hoc exploration of the BC crime dataset.
Copy and modify the code blocks below as building blocks for your own analyses.

# Setup

```{r load-packages}
packages <- c("dplyr", "ggplot2", "tidyr", "scales", "knitr", "glue")

installed_packages <- packages %in% rownames(installed.packages())
if (any(!installed_packages)) {
  install.packages(packages[!installed_packages])
}

library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(knitr)
```

```{r load-data}
csv_path <- "../data/bc_crime.csv"
if (!file.exists(csv_path)) {
  stop("Data file not found. Place bc_crime.csv in the data/ directory.")
}
crime_data <- read.csv(csv_path, stringsAsFactors = FALSE)
```

```{r theme}
# Matching the main report's visual identity
bc_accent <- "#08519c"
bc_red    <- "#c0392b"
bc_green  <- "#27ae60"

theme_bc_crime <- function(base_size = 13) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      plot.title = element_text(face = "bold", size = rel(1.25), hjust = 0),
      plot.subtitle = element_text(color = "grey40", size = rel(0.9)),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank()
    )
}
```

# Quick Summary

```{r summary-overview, results='asis'}
cat(glue::glue(
  "- **Records:** {comma(nrow(crime_data))}\n",
  "- **Years:** {min(crime_data$YEAR)} to {max(crime_data$YEAR)}\n",
  "- **Crime types:** {length(unique(crime_data$TYPE))}\n",
  "- **Neighbourhoods:** {length(unique(crime_data$NEIGHBOURHOOD))}\n"
))
```

```{r preview}
kable(head(crime_data, 10), caption = "First 10 records")
```

# Exploration Templates

## Template 1: Crime counts by year

```{r yearly-counts}
crime_data %>%
  count(YEAR) %>%
  ggplot(aes(x = YEAR, y = n)) +
  geom_col(fill = bc_accent, width = 0.6) +
  geom_text(aes(label = comma(n)), vjust = -0.5, size = 3.5) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.12))) +
  scale_x_continuous(breaks = unique(crime_data$YEAR)) +
  labs(title = "Crime Count per Year", x = "Year", y = "Count") +
  theme_bc_crime()
```

## Template 2: Filter by crime type

Change `"Homicide"` to any crime type in the data.

```{r type-filter}
TARGET_TYPE <- "Homicide"  # <-- Change this

crime_data %>%
  filter(TYPE == TARGET_TYPE) %>%
  count(YEAR) %>%
  ggplot(aes(x = YEAR, y = n)) +
  geom_col(fill = bc_red, width = 0.6) +
  geom_text(aes(label = n), vjust = -0.5, fontface = "bold") +
  scale_x_continuous(breaks = unique(crime_data$YEAR)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = paste(TARGET_TYPE, "Count per Year"), x = "Year", y = "Count") +
  theme_bc_crime()
```

## Template 3: Neighbourhood comparison (single year)

```{r neighbourhood-template}
TARGET_YEAR <- max(crime_data$YEAR)  # <-- Change this
TOP_N <- 10

crime_data %>%
  filter(YEAR == TARGET_YEAR, NEIGHBOURHOOD != "Central Business District") %>%
  count(NEIGHBOURHOOD, sort = TRUE) %>%
  head(TOP_N) %>%
  ggplot(aes(x = reorder(NEIGHBOURHOOD, n), y = n, fill = n)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_viridis_c(option = "D", direction = -1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = paste("Top", TOP_N, "Neighbourhoods by Crime Volume (", TARGET_YEAR, ")"),
    x = NULL, y = "Total Crimes"
  ) +
  theme_bc_crime()
```

## Template 4: Cross-tabulation

```{r cross-tab}
crime_data %>%
  filter(YEAR == max(YEAR)) %>%
  count(TYPE, sort = TRUE) %>%
  mutate(
    share = round(n / sum(n) * 100, 1),
    cumulative = round(cumsum(n) / sum(n) * 100, 1)
  ) %>%
  kable(
    caption = paste("Crime type distribution (", max(crime_data$YEAR), ")"),
    format.args = list(big.mark = ",")
  )
```

## Template 5: Year-over-year change

```{r yoy-template}
crime_data %>%
  count(YEAR) %>%
  mutate(change_pct = round((n / lag(n) - 1) * 100, 1)) %>%
  filter(!is.na(change_pct)) %>%
  ggplot(aes(x = factor(YEAR), y = change_pct, fill = change_pct > 0)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.6) +
  geom_text(aes(label = paste0(ifelse(change_pct > 0, "+", ""), change_pct, "%")),
            vjust = ifelse(crime_data %>% count(YEAR) %>% mutate(c = round((n / lag(n) - 1) * 100, 1)) %>% filter(!is.na(c)) %>% pull(c) > 0, -0.5, 1.5),
            size = 3.5, fontface = "bold") +
  scale_fill_manual(values = c("TRUE" = bc_red, "FALSE" = bc_green)) +
  labs(title = "Year-over-Year Crime Change", x = NULL, y = "Change (%)") +
  theme_bc_crime()
```

---

*Use this worksheet as a starting point. Copy and modify the templates above
for your own investigations. See `BC-CRIME.Rmd` for the full analysis.*
