---
title: "Exploring the Crime Landscape in British Columbia, Canada"
author: "Casper Kacper Ruta"
date: "`r Sys.Date()`"
params:
  data_file: "../data/bc_crime.csv"
  exclude_neighbourhoods: "Central Business District"
  top_n_types: 8
  top_n_hoods: 10
  forecast_years: 2
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    theme: flatly
    highlight: tango
    code_folding: hide
    number_sections: true
    fig_width: 10
    fig_height: 6
    fig_caption: true
    df_print: kable
    css: custom.css
  pdf_document:
    toc: true
    fig_width: 10
    fig_height: 6
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.retina = 2,
  out.width = "100%"
)
```

# Abstract {-}

```{r abstract, echo=FALSE, results='asis'}
cat(paste0(
  "This report analyses ", comma(nrow(crime_data)), " crime records across British Columbia, ",
  "Canada, spanning ", min(crime_data$YEAR), " to ", max(crime_data$YEAR), ". ",
  "The analysis examines trends in total crime volume, compositional shifts across ",
  length(unique(crime_data$TYPE)), " offence categories, geographic concentration ",
  "patterns across ", length(unique(crime_data$NEIGHBOURHOOD)), " neighbourhoods, ",
  "and seasonal variation. Key findings include: (1) the overall trend direction ",
  "and its statistical significance, (2) which crime types are driving aggregate changes, ",
  "(3) how crime is geographically concentrated and which areas are improving or worsening, ",
  "and (4) forward-looking projections based on current trajectories. ",
  "Methods include descriptive statistics, chi-squared distributional tests, ",
  "hierarchical clustering of neighbourhood crime profiles, ",
  "linear trend regression with confidence intervals, and severity-weighted indexing. ",
  "All analyses are fully reproducible from the source R Markdown file."
))
```

---

# Setup and Data Loading {-}

```{r load-packages}
packages <- c("here", "readr", "stringr", "clock", "dplyr", "ggplot2",
              "rmarkdown", "knitr", "magrittr", "glue", "xfun", "fs",
              "tidyr", "scales", "DT")

installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(knitr)
```

```{r load-data}
csv_path <- params$data_file
if (!file.exists(csv_path)) {
  stop(paste0(
    "Data file not found: ", csv_path, "\n",
    "Please place bc_crime.csv in the data/ directory.\n",
    "Expected columns: YEAR, TYPE, NEIGHBOURHOOD (MONTH optional)."
  ))
}

crime_data <- tryCatch(
  read.csv(csv_path, stringsAsFactors = FALSE),
  error = function(e) stop(paste("Failed to read CSV:", e$message))
)

# Validate required columns
required_cols <- c("YEAR", "TYPE", "NEIGHBOURHOOD")
missing_cols <- setdiff(required_cols, names(crime_data))
if (length(missing_cols) > 0) {
  stop(paste("Missing required columns:", paste(missing_cols, collapse = ", ")))
}

cat(glue::glue("Loaded {comma(nrow(crime_data))} records with {ncol(crime_data)} columns."), "\n")
```

```{r config}
# ------------------------------------------------------------------
# Report Configuration
# Values sourced from YAML params; override via rmarkdown::render()
# Example: rmarkdown::render("BC-CRIME.Rmd", params = list(top_n_hoods = 15))
# ------------------------------------------------------------------
LATEST_YEAR     <- max(crime_data$YEAR)
EARLIEST_YEAR   <- min(crime_data$YEAR)
COMPARE_YEARS   <- c(EARLIEST_YEAR, EARLIEST_YEAR + 1, LATEST_YEAR)
EXCLUDE_HOODS   <- strsplit(params$exclude_neighbourhoods, ",\\s*")[[1]]
TOP_N_TYPES     <- params$top_n_types
TOP_N_HOODS     <- params$top_n_hoods
TOP_N_HOODS_PROFILE <- 8
TOP_N_TYPES_PROFILE <- 6
TOP_N_DUMBBELL  <- 12
COVID_START     <- 2019.5
COVID_END       <- 2020.5
has_month       <- "MONTH" %in% names(crime_data)
has_DT          <- requireNamespace("DT", quietly = TRUE)
```

```{r custom-theme}
# ------------------------------------------------------------------
# Unified visual identity for all charts
# ------------------------------------------------------------------
theme_bc_crime <- function(base_size = 13) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      plot.title       = element_text(face = "bold", size = rel(1.25), hjust = 0,
                                      margin = margin(b = 4)),
      plot.subtitle    = element_text(color = "grey40", size = rel(0.9), hjust = 0,
                                      margin = margin(b = 10)),
      plot.caption     = element_text(color = "grey60", size = rel(0.7), hjust = 1,
                                      margin = margin(t = 8)),
      axis.title       = element_text(face = "bold", size = rel(0.85)),
      axis.text        = element_text(size = rel(0.8)),
      legend.position  = "bottom",
      legend.title     = element_text(face = "bold", size = rel(0.8)),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      strip.text       = element_text(face = "bold", size = rel(0.9)),
      strip.background = element_rect(fill = "grey95", color = NA),
      plot.margin      = margin(10, 15, 10, 10)
    )
}

# Palettes
bc_accent   <- "#08519c"
bc_red      <- "#c0392b"
bc_green    <- "#27ae60"
bc_palette  <- c("#08519c", "#3182bd", "#6baed6", "#9ecae1", "#c6dbef", "#deebf7")

# Helper: safe division avoiding Inf / NaN
safe_pct_change <- function(new, old) {
  ifelse(old == 0, NA_real_, round((new / old - 1) * 100, 1))
}

# Reusable horizontal bar chart for neighbourhood data
plot_hood_bar <- function(data, title, subtitle = NULL, show_labels = TRUE) {
  p <- ggplot(data, aes(x = reorder(NEIGHBOURHOOD, n), y = n, fill = n)) +
    geom_col(show.legend = FALSE) +
    coord_flip() +
    scale_fill_viridis_c(option = "D", direction = -1) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
    labs(title = title, subtitle = subtitle, x = NULL, y = "Total Reported Crimes",
         caption = "Source: BC municipal crime data") +
    theme_bc_crime()
  if (show_labels) {
    p <- p + geom_text(aes(label = comma(n)), hjust = -0.05, size = 3.2, fontface = "bold")
  }
  p
}

# Auto-incrementing figure counter for cross-referencing
fig_counter <- local({
  count <- 0
  function(caption) {
    count <<- count + 1
    glue::glue("Figure {count}: {caption}")
  }
})

# Auto-incrementing table counter
tbl_counter <- local({
  count <- 0
  function(caption) {
    count <<- count + 1
    glue::glue("Table {count}: {caption}")
  }
})

# Helper: get top N values from a column, returns a character vector
get_top_n <- function(data, col, n = 10) {
  data %>%
    count({{ col }}, sort = TRUE) %>%
    head(n) %>%
    pull({{ col }})
}

# Helper: add COVID-19 annotation band to a ggplot
add_covid_band <- function(plot, y_max = NULL) {
  plot +
    annotate("rect", xmin = COVID_START, xmax = COVID_END,
             ymin = -Inf, ymax = Inf, fill = bc_red, alpha = 0.06) +
    annotate("text", x = mean(c(COVID_START, COVID_END)),
             y = if (!is.null(y_max)) y_max * 1.08 else Inf,
             label = "COVID-19\nonset", color = bc_red, size = 3,
             fontface = "italic", vjust = if (is.null(y_max)) 1.2 else 0)
}

# Helper: format percentage change with directional arrow
format_change <- function(value) {
  ifelse(is.na(value), "N/A",
    ifelse(value > 0, paste0("\u2191 +", value, "%"),
      ifelse(value < 0, paste0("\u2193 ", value, "%"),
        "0%")))
}

# Helper: create a summary stat for inline reporting
stat_inline <- function(x, fmt = comma) {
  fmt(x)
}
```

# Executive Summary

```{r exec-summary, echo=FALSE, results='asis'}
exec_total   <- nrow(crime_data)
exec_years   <- range(crime_data$YEAR)
exec_latest  <- crime_data %>% filter(YEAR == max(YEAR)) %>% nrow()
exec_first   <- crime_data %>% filter(YEAR == min(YEAR)) %>% nrow()
exec_change  <- safe_pct_change(exec_latest, exec_first)
exec_top     <- crime_data %>% filter(YEAR == max(YEAR)) %>% count(TYPE, sort = TRUE) %>% slice(1)
exec_n_types <- length(unique(crime_data$TYPE))
exec_n_hoods <- length(unique(crime_data$NEIGHBOURHOOD))

change_color <- ifelse(exec_change > 0, "#c0392b", "#27ae60")
change_sign  <- ifelse(exec_change > 0, "+", "")

cat(glue::glue('
<div style="display: flex; flex-wrap: wrap; gap: 12px; margin: 1rem 0;">
<div style="flex: 1; min-width: 140px; background: #f0f4f8; border-left: 4px solid #08519c; padding: 12px 16px; border-radius: 0 6px 6px 0;">
<div style="font-size: 0.8em; color: #666; text-transform: uppercase; letter-spacing: 0.5px;">Total Records</div>
<div style="font-size: 1.5em; font-weight: bold; color: #08519c;">{comma(exec_total)}</div>
</div>
<div style="flex: 1; min-width: 140px; background: #f0f4f8; border-left: 4px solid #08519c; padding: 12px 16px; border-radius: 0 6px 6px 0;">
<div style="font-size: 0.8em; color: #666; text-transform: uppercase; letter-spacing: 0.5px;">Period</div>
<div style="font-size: 1.5em; font-weight: bold; color: #08519c;">{exec_years[1]}&ndash;{exec_years[2]}</div>
</div>
<div style="flex: 1; min-width: 140px; background: #f0f4f8; border-left: 4px solid {change_color}; padding: 12px 16px; border-radius: 0 6px 6px 0;">
<div style="font-size: 0.8em; color: #666; text-transform: uppercase; letter-spacing: 0.5px;">Overall Change</div>
<div style="font-size: 1.5em; font-weight: bold; color: {change_color};">{change_sign}{exec_change}%</div>
</div>
<div style="flex: 1; min-width: 140px; background: #f0f4f8; border-left: 4px solid #08519c; padding: 12px 16px; border-radius: 0 6px 6px 0;">
<div style="font-size: 0.8em; color: #666; text-transform: uppercase; letter-spacing: 0.5px;">Top Crime ({exec_years[2]})</div>
<div style="font-size: 1.1em; font-weight: bold; color: #08519c;">{exec_top$TYPE}</div>
<div style="font-size: 0.9em; color: #888;">{comma(exec_top$n)} incidents</div>
</div>
</div>

| Metric | Value |
|:-------|------:|
| **Crime types** | {exec_n_types} |
| **Neighbourhoods** | {exec_n_hoods} |
| **Latest year volume** | {comma(exec_latest)} |
'))
```

# Introduction

This report examines crime trends across British Columbia, Canada, using publicly
available municipal crime data. The analysis spans **`r exec_years[1]`** to
**`r exec_years[2]`** and is organized into six parts:

1. **Overall Trends** --- aggregate crime volume and year-over-year momentum
2. **Crime Type Analysis** --- compositional breakdown by offence category
3. **Neighbourhood Analysis** --- geographic variation, clustering, and hotspot identification
4. **Seasonal & Monthly Patterns** --- temporal rhythms within each year
5. **Crime Severity Index** --- a weighted view beyond raw counts
6. **Trend Projections** --- forward-looking estimates with uncertainty bounds

The overarching question: *How has crime in British Columbia evolved in the years
surrounding the COVID-19 pandemic, and which communities and crime types are driving
those changes?*

## Broader Context

Understanding local crime data requires broader context. Key reference points:

- **Population growth:** British Columbia's population grew by approximately 7.6%
  between 2018 and 2023 (from ~5.0M to ~5.4M), meaning *per-capita* crime rates
  declined even if absolute counts remained flat. Population-normalized analysis
  would require neighbourhood-level population estimates not available in this dataset.
- **National trends:** Statistics Canada's Crime Severity Index (CSI) for BC has
  historically tracked above the national average but followed similar pandemic-era
  patterns: a dip in 2020 followed by a partial rebound.
- **COVID-19 impacts:** The pandemic fundamentally altered crime patterns across Canada.
  Property crime fell (fewer targets available during lockdowns), while some personal
  offences (domestic violence) increased. These national patterns provide a baseline
  for interpreting local BC data.
- **Policy context:** BC has implemented several policing reforms during this period,
  including increased emphasis on restorative justice and community policing models,
  which may influence both crime rates and reporting behaviour.

> **Analytical limitation:** This report analyses *reported* crime counts. Changes
> in reporting rates, police staffing, or recording practices can affect observed
> trends independently of actual crime levels.

---

# Data Quality Check

Before analysis, it is good practice to validate the dataset for completeness
and consistency. This section runs automated checks.

```{r data-quality, results='asis'}
# Automated data quality checks
checks <- list()

# Check 1: Missing values
na_counts <- colSums(is.na(crime_data))
na_total  <- sum(na_counts)
checks$missing <- ifelse(na_total == 0,
  "**PASS** --- No missing values detected.",
  glue::glue("**WARNING** --- {na_total} missing values found: {paste(names(na_counts[na_counts > 0]), collapse = ', ')}"))

# Check 2: Year continuity
years <- sort(unique(crime_data$YEAR))
year_gaps <- setdiff(seq(min(years), max(years)), years)
checks$year_gaps <- ifelse(length(year_gaps) == 0,
  glue::glue("**PASS** --- Continuous year coverage ({min(years)}\u2013{max(years)})."),
  glue::glue("**WARNING** --- Missing years: {paste(year_gaps, collapse = ', ')}"))

# Check 3: Record count per year (check for suspiciously low years)
year_counts <- crime_data %>% count(YEAR)
min_year <- year_counts %>% filter(n == min(n))
max_year <- year_counts %>% filter(n == max(n))
ratio <- round(max_year$n[1] / min_year$n[1], 1)
checks$balance <- glue::glue(
  "**INFO** --- Year with fewest records: {min_year$YEAR[1]} ({comma(min_year$n[1])}). ",
  "Year with most: {max_year$YEAR[1]} ({comma(max_year$n[1])}). Ratio: {ratio}x.")

# Check 4: Duplicate records
n_dupes <- sum(duplicated(crime_data))
checks$duplicates <- ifelse(n_dupes == 0,
  "**PASS** --- No exact duplicate rows detected.",
  glue::glue("**WARNING** --- {comma(n_dupes)} duplicate rows found ({round(n_dupes/nrow(crime_data)*100, 2)}% of data)."))

# Check 5: Crime type consistency across years
type_by_year <- crime_data %>% group_by(YEAR) %>% summarise(n_types = n_distinct(TYPE))
type_range <- range(type_by_year$n_types)
checks$type_consistency <- ifelse(type_range[1] == type_range[2],
  glue::glue("**PASS** --- All years have exactly {type_range[1]} crime types."),
  glue::glue("**INFO** --- Crime types per year range from {type_range[1]} to {type_range[2]}. ",
             "New types may have been added or removed over time."))

# Check 6: Neighbourhood consistency
hood_by_year <- crime_data %>% group_by(YEAR) %>% summarise(n_hoods = n_distinct(NEIGHBOURHOOD))
hood_range <- range(hood_by_year$n_hoods)
checks$hood_consistency <- ifelse(hood_range[1] == hood_range[2],
  glue::glue("**PASS** --- All years have exactly {hood_range[1]} neighbourhoods."),
  glue::glue("**INFO** --- Neighbourhoods per year range from {hood_range[1]} to {hood_range[2]}."))

# Check 7: Outlier detection (years with unusual volume using IQR)
q1 <- quantile(year_counts$n, 0.25)
q3 <- quantile(year_counts$n, 0.75)
iqr_val <- q3 - q1
outlier_years <- year_counts %>% filter(n < q1 - 1.5 * iqr_val | n > q3 + 1.5 * iqr_val)
checks$outliers <- ifelse(nrow(outlier_years) == 0,
  "**PASS** --- No outlier years detected (IQR method).",
  glue::glue("**INFO** --- Potential outlier year(s): {paste(outlier_years$YEAR, collapse = ', ')} ",
             "(volume outside 1.5x IQR)."))

# Check 8: Column completeness per column
col_completeness <- sapply(crime_data, function(x) round(sum(!is.na(x)) / length(x) * 100, 1))
low_cols <- names(col_completeness[col_completeness < 100])
checks$completeness <- ifelse(length(low_cols) == 0,
  "**PASS** --- All columns 100% complete.",
  glue::glue("**WARNING** --- Columns with missing data: {paste(low_cols, collapse = ', ')}."))

cat("| Check | Result |\n|:------|:-------|\n")
cat(glue::glue("| Missing values | {checks$missing} |\n"))
cat(glue::glue("| Year continuity | {checks$year_gaps} |\n"))
cat(glue::glue("| Record balance | {checks$balance} |\n"))
cat(glue::glue("| Duplicate rows | {checks$duplicates} |\n"))
cat(glue::glue("| Crime type consistency | {checks$type_consistency} |\n"))
cat(glue::glue("| Neighbourhood consistency | {checks$hood_consistency} |\n"))
cat(glue::glue("| Volume outliers | {checks$outliers} |\n"))
cat(glue::glue("| Column completeness | {checks$completeness} |\n"))
```

---

# Data Overview

```{r data-overview, results='asis'}
n_records <- format(nrow(crime_data), big.mark = ",")
n_types   <- length(unique(crime_data$TYPE))
n_hoods   <- length(unique(crime_data$NEIGHBOURHOOD))
yr_range  <- range(crime_data$YEAR)

cat(glue::glue(
  "The dataset contains **{n_records}** individual crime records spanning ",
  "**{yr_range[1]}** to **{yr_range[2]}**, encompassing **{n_types}** offence ",
  "categories across **{n_hoods}** neighbourhoods.\n\n"
))
```

```{r data-preview}
kable(
  head(crime_data, 8),
  caption = "Table 1: First 8 records of the BC crime dataset"
)
```

```{r data-column-summary}
col_summary <- data.frame(
  Column = names(crime_data),
  Class  = sapply(crime_data, class),
  Unique = sapply(crime_data, function(x) length(unique(x))),
  Missing = sapply(crime_data, function(x) sum(is.na(x))),
  row.names = NULL
)
kable(col_summary, caption = "Table 2: Column-level summary statistics")
```

---

# Part 1: Overall Crime Trends

This section tracks aggregate crime volume over time. Understanding the macro trend
is essential before drilling into specific crime types or neighbourhoods.

## Annual Crime Volume

```{r annual-crime-trend, fig.alt="Line chart showing total reported crimes by year with COVID-19 annotation band"}
crime_by_year <- crime_data %>%
  count(YEAR) %>%
  mutate(
    pct_change = safe_pct_change(n, lag(n)),
    label = comma(n),
    cumulative = cumsum(n)
  )

ggplot(crime_by_year, aes(x = YEAR, y = n)) +
  annotate("rect", xmin = COVID_START, xmax = COVID_END,
           ymin = -Inf, ymax = Inf, fill = bc_red, alpha = 0.06) +
  annotate("text", x = 2020, y = max(crime_by_year$n) * 1.08,
           label = "COVID-19\nonset", color = bc_red, size = 3, fontface = "italic") +
  geom_ribbon(aes(ymin = min(n) * 0.95, ymax = n), fill = bc_accent, alpha = 0.08) +
  geom_line(color = bc_accent, linewidth = 1.3) +
  geom_point(color = bc_accent, size = 3.5, shape = 21, fill = "white", stroke = 1.5) +
  geom_text(aes(label = label), vjust = -1.4, size = 3.5, color = "grey25") +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0.03, 0.18))) +
  scale_x_continuous(breaks = crime_by_year$YEAR) +
  labs(
    title = "Total Reported Crimes by Year",
    subtitle = glue::glue("British Columbia, {yr_range[1]}\u2013{yr_range[2]} | Shaded band marks the COVID-19 pandemic year"),
    x = NULL, y = "Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r annual-change-table}
kable(
  crime_by_year %>%
    select(YEAR, n, pct_change) %>%
    rename(Year = YEAR, `Total Crimes` = n, `YoY Change (%)` = pct_change),
  caption = "Table 3: Annual crime counts with year-over-year percentage change",
  format.args = list(big.mark = ",")
)
```

```{r trend-interpretation, results='asis'}
# Linear trend with statistical significance
trend_model <- lm(n ~ YEAR, data = crime_by_year)
trend_summary <- summary(trend_model)
slope <- coef(trend_model)[["YEAR"]]
slope_dir <- ifelse(slope > 0, "increasing", "decreasing")
slope_per_year <- round(abs(slope), 0)
p_value <- trend_summary$coefficients["YEAR", "Pr(>|t|)"]
r_squared <- round(trend_summary$r.squared, 3)
sig_label <- ifelse(p_value < 0.01, "highly significant (p < 0.01)",
              ifelse(p_value < 0.05, "statistically significant (p < 0.05)",
                ifelse(p_value < 0.10, "marginally significant (p < 0.10)",
                  "not statistically significant (p \u2265 0.10)")))

# Confidence interval on slope
ci <- confint(trend_model, "YEAR", level = 0.95)
ci_low <- round(abs(ci[1]), 0)
ci_high <- round(abs(ci[2]), 0)

cat(glue::glue(
  "> **Trend summary:** A linear regression finds reported crimes are ",
  "**{slope_dir}** by approximately **{comma(slope_per_year)} incidents per year** ",
  "(95% CI: {comma(min(ci_low, ci_high))} to {comma(max(ci_low, ci_high))}). ",
  "This trend is **{sig_label}** with R\u00b2 = {r_squared}, explaining ",
  "{round(r_squared * 100, 1)}% of the year-to-year variance. ",
  "Note: the linear approximation does not capture non-linear pandemic effects.\n\n"
))
```

## Year-over-Year Momentum

The bar chart below isolates the *direction and magnitude* of change each year.
Green bars indicate a year where crime fell relative to the prior year; red bars
indicate an increase.

```{r yoy-bar, fig.alt="Bar chart showing year-over-year percentage change in crime, red for increases, green for decreases"}
yoy_data <- crime_by_year %>% filter(!is.na(pct_change))

ggplot(yoy_data, aes(x = factor(YEAR), y = pct_change, fill = pct_change > 0)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.6, color = "grey40") +
  geom_text(
    aes(label = paste0(ifelse(pct_change > 0, "+", ""), pct_change, "%")),
    vjust = ifelse(yoy_data$pct_change > 0, -0.6, 1.6),
    size = 4, fontface = "bold"
  ) +
  scale_fill_manual(values = c("TRUE" = bc_red, "FALSE" = bc_green)) +
  scale_y_continuous(labels = function(x) paste0(x, "%"),
                     expand = expansion(mult = 0.2)) +
  labs(
    title = "Year-over-Year Change in Total Reported Crimes",
    subtitle = "Red = increase over prior year | Green = decrease",
    x = NULL, y = "Change (%)",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r part1-takeaway, results='asis', echo=FALSE}
latest_change <- tail(crime_by_year$pct_change, 1)
cat(glue::glue(
  "\n> **Part 1 takeaway:** The aggregate trend is **{slope_dir}** at ~{comma(slope_per_year)} ",
  "incidents/year. The most recent year-over-year change was **{format_change(latest_change)}**. ",
  "But aggregate trends can mask divergent patterns across crime types and areas---the next ",
  "sections unpack this.\n\n"
))
```

---

# Part 2: Crime Type Analysis {.tabset .tabset-fade}

Not all crimes are equal in severity or frequency. This section decomposes the
aggregate trend by offence type to understand *what kinds* of crime are rising or
falling. Use the tabs below to explore different angles.

## By Count

```{r crime-type-table}
crime_counts <- crime_data %>%
  group_by(YEAR, TYPE) %>%
  summarise(Count = n(), .groups = "drop")

crime_type_wide <- crime_counts %>%
  pivot_wider(names_from = YEAR, values_from = Count, values_fill = 0)

kable(
  crime_type_wide,
  caption = glue::glue("Table 4: Crime counts by type and year ({yr_range[1]}\u2013{yr_range[2]})"),
  format.args = list(big.mark = ",")
)
```

```{r chi-squared-test, results='asis'}
# Chi-squared test: Has the distribution of crime types changed significantly?
contingency <- crime_data %>%
  filter(YEAR %in% c(EARLIEST_YEAR, LATEST_YEAR)) %>%
  count(YEAR, TYPE) %>%
  pivot_wider(names_from = YEAR, values_from = n, values_fill = 0)

chi_mat <- as.matrix(contingency[, -1])
rownames(chi_mat) <- contingency$TYPE
chi_test <- chisq.test(chi_mat)
chi_p <- chi_test$p.value
chi_sig <- ifelse(chi_p < 0.001, "highly significant (p < 0.001)",
            ifelse(chi_p < 0.05, paste0("significant (p = ", round(chi_p, 4), ")"),
              paste0("not significant (p = ", round(chi_p, 3), ")")))

cat(glue::glue(
  "> **Distribution test:** A chi-squared test comparing crime type distributions ",
  "between {EARLIEST_YEAR} and {LATEST_YEAR} is **{chi_sig}** ",
  "(\u03c7\u00b2 = {round(chi_test$statistic, 1)}, df = {chi_test$parameter}), ",
  "indicating that the *mix* of crime types has ",
  "{ifelse(chi_p < 0.05, 'shifted meaningfully', 'remained largely stable')} ",
  "over this period.\n\n"
))
```

## Distribution

Before examining trends, here is the overall breakdown for the most recent year.
This proportional bar chart replaces a pie chart --- it is easier to compare
relative sizes accurately.

```{r crime-type-proportion}
type_dist <- crime_data %>%
  filter(YEAR == max(YEAR)) %>%
  count(TYPE, sort = TRUE) %>%
  mutate(
    pct = round(n / sum(n) * 100, 1),
    label = ifelse(pct >= 2, paste0(TYPE, "\n", pct, "%"), "")
  )

ggplot(type_dist, aes(x = 1, y = n, fill = reorder(TYPE, -n))) +
  geom_col(width = 1, color = "white", linewidth = 0.3, show.legend = TRUE) +
  coord_flip() +
  scale_fill_brewer(palette = "Set3", name = "Crime Type") +
  labs(
    title = glue::glue("Crime Type Distribution ({max(crime_data$YEAR)})"),
    subtitle = "Proportional bar; each segment represents one crime type's share",
    x = NULL, y = NULL,
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.position = "right"
  )
```

## Composition

The stacked area chart reveals how the *mix* of crime types shifts across years.
The top 8 categories are shown individually; less common types are grouped.

```{r crime-type-area, fig.height=7}
top_types <- crime_data %>%
  count(TYPE, sort = TRUE) %>%
  head(8) %>%
  pull(TYPE)

crime_type_trend <- crime_data %>%
  mutate(TYPE_grouped = ifelse(TYPE %in% top_types, TYPE, "Other")) %>%
  count(YEAR, TYPE_grouped) %>%
  rename(TYPE = TYPE_grouped)

ggplot(crime_type_trend, aes(x = YEAR, y = n, fill = TYPE)) +
  geom_area(alpha = 0.85, color = "white", linewidth = 0.3) +
  scale_fill_brewer(palette = "Set2", name = "Crime Type") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = unique(crime_type_trend$YEAR)) +
  labs(
    title = "Crime Composition Over Time",
    subtitle = "Top 8 crime types shown individually; remaining grouped as 'Other'",
    x = NULL, y = "Number of Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right")
```

## Heatmap

This heatmap normalizes each year's data to show the *percentage share* of each
crime type, making it easy to spot shifts in crime composition independent of
overall volume changes.

```{r crime-heatmap, fig.height=9, fig.width=10, fig.alt="Heatmap of crime type proportions by year"}
crime_type_year <- crime_data %>%
  count(YEAR, TYPE) %>%
  group_by(YEAR) %>%
  mutate(proportion = round(n / sum(n) * 100, 1)) %>%
  ungroup()

# Order types by total volume for better readability
type_order <- crime_type_year %>%
  group_by(TYPE) %>%
  summarise(total = sum(n)) %>%
  arrange(total) %>%
  pull(TYPE)

crime_type_year <- crime_type_year %>%
  mutate(TYPE = factor(TYPE, levels = type_order))

ggplot(crime_type_year, aes(x = factor(YEAR), y = TYPE, fill = proportion)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(aes(label = ifelse(proportion >= 0.5, paste0(proportion, "%"), ""),
                color = proportion > median(proportion)),
            size = 3, fontface = "bold", show.legend = FALSE) +
  scale_color_manual(values = c("TRUE" = "white", "FALSE" = "grey20")) +
  scale_fill_viridis_c(option = "C", name = "Share of\nYear's Total (%)") +
  labs(
    title = "Crime Type Proportions by Year",
    subtitle = "Each cell shows the percentage share within that year; types ordered by total volume",
    x = "Year", y = NULL,
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right", panel.grid = element_blank())
```

## Homicides

Homicides, while a small fraction of total crime, represent the most severe
category. The dedicated chart below tracks their year-over-year trajectory.

```{r homicide-trend}
homicide_count <- crime_data %>%
  filter(TYPE == "Homicide") %>%
  count(YEAR) %>%
  mutate(
    pct_change = safe_pct_change(n, lag(n)),
    avg = mean(n)
  )

ggplot(homicide_count, aes(x = YEAR, y = n)) +
  geom_hline(aes(yintercept = avg), linetype = "dashed", color = "grey50", linewidth = 0.5) +
  annotate("text", x = min(homicide_count$YEAR) + 0.3, y = homicide_count$avg[1] + 0.5,
           label = paste0("Average: ", round(homicide_count$avg[1], 1)),
           color = "grey50", size = 3.5, hjust = 0) +
  geom_col(fill = bc_red, width = 0.55, alpha = 0.85) +
  geom_text(aes(label = n), vjust = -0.6, fontface = "bold", size = 4.2) +
  scale_x_continuous(breaks = homicide_count$YEAR) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(
    title = "Homicide Count by Year",
    subtitle = "Dashed line = period average across all years",
    x = NULL, y = "Number of Homicides",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r homicide-table}
kable(
  homicide_count %>%
    select(YEAR, n, pct_change) %>%
    rename(Year = YEAR, Homicides = n, `YoY Change (%)` = pct_change),
  caption = "Table 5: Annual homicide counts with year-over-year change"
)
```

## Rankings

This bump chart shows how the relative ranking of crime types has shifted. A rising
line means a crime type is becoming relatively more common.

```{r rank-bump, fig.height=7}
type_ranks <- crime_data %>%
  count(YEAR, TYPE) %>%
  group_by(YEAR) %>%
  mutate(rank = dense_rank(desc(n))) %>%
  ungroup() %>%
  filter(TYPE %in% top_types)

ggplot(type_ranks, aes(x = YEAR, y = rank, color = TYPE, group = TYPE)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2.5) +
  scale_y_reverse(breaks = 1:max(type_ranks$rank)) +
  scale_x_continuous(breaks = unique(type_ranks$YEAR)) +
  scale_color_brewer(palette = "Set2", name = "Crime Type") +
  labs(
    title = "Crime Type Rankings Over Time",
    subtitle = "Top 8 crime types by volume; rank 1 = most common that year",
    x = NULL, y = "Rank (1 = most common)",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right", panel.grid.major.x = element_line(color = "grey90"))
```

## Individual Trends

Each panel below tracks one crime type independently, making it easy to see which
types are rising, falling, or volatile year to year.

```{r small-multiples, fig.height=10, fig.width=12}
type_trends <- crime_data %>%
  count(YEAR, TYPE) %>%
  group_by(TYPE) %>%
  mutate(
    first_val = first(n),
    indexed = round(n / first_val * 100, 0)
  ) %>%
  ungroup()

ggplot(type_trends, aes(x = YEAR, y = n)) +
  geom_line(color = bc_accent, linewidth = 0.8) +
  geom_point(color = bc_accent, size = 1.5) +
  facet_wrap(~ TYPE, scales = "free_y", ncol = 4) +
  scale_x_continuous(breaks = c(min(type_trends$YEAR), max(type_trends$YEAR))) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Individual Crime Type Trends",
    subtitle = "Each panel uses its own y-axis scale to highlight per-type patterns",
    x = NULL, y = NULL,
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime(base_size = 10) +
  theme(
    strip.text = element_text(size = 8),
    panel.grid.major.x = element_line(color = "grey92")
  )
```

## Type Correlations

Do different crime types move together or in opposite directions? This correlation
matrix shows which crime types tend to increase in the same years.

```{r type-correlation, fig.height=7, fig.width=8}
type_cor_data <- crime_data %>%
  filter(TYPE %in% top_types) %>%
  count(YEAR, TYPE) %>%
  pivot_wider(names_from = TYPE, values_from = n, values_fill = 0)

# Compute correlation matrix (exclude YEAR column)
cor_matrix <- cor(type_cor_data[, -1])

# Convert to long format for ggplot
cor_long <- as.data.frame(as.table(cor_matrix)) %>%
  rename(Type1 = Var1, Type2 = Var2, Correlation = Freq)

ggplot(cor_long, aes(x = Type1, y = Type2, fill = Correlation)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = round(Correlation, 2),
                color = abs(Correlation) > 0.5),
            size = 3, fontface = "bold", show.legend = FALSE) +
  scale_color_manual(values = c("TRUE" = "white", "FALSE" = "grey20")) +
  scale_fill_gradient2(low = bc_green, mid = "white", high = bc_red,
                       midpoint = 0, limits = c(-1, 1), name = "Correlation") +
  labs(
    title = "Crime Type Correlation Matrix",
    subtitle = "Do crime types rise and fall together? Red = positive correlation, Green = negative",
    x = NULL, y = NULL,
    caption = "Based on annual counts. Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 40, hjust = 1, size = 8),
    axis.text.y = element_text(size = 8),
    legend.position = "right"
  )
```

```{r cor-interpretation, results='asis'}
# Find the strongest positive and negative correlations (excluding self-correlations)
cor_filtered <- cor_long %>%
  filter(as.character(Type1) < as.character(Type2)) %>%
  arrange(desc(Correlation))

strongest_pos <- cor_filtered %>% slice(1)
strongest_neg <- cor_filtered %>% arrange(Correlation) %>% slice(1)

cat(glue::glue(
  "> **Correlation insight:** The strongest positive correlation is between ",
  "*{strongest_pos$Type1}* and *{strongest_pos$Type2}* (r = {round(strongest_pos$Correlation, 2)}), ",
  "suggesting these crimes move in tandem. ",
  "The weakest correlation involves *{strongest_neg$Type1}* and *{strongest_neg$Type2}* ",
  "(r = {round(strongest_neg$Correlation, 2)}).\n\n"
))
```

## Indexed Comparison

Comparing crime types of very different magnitudes (e.g., thousands of thefts vs
dozens of homicides) on the same scale is misleading. By indexing each type to
100 in the first available year, we can compare *relative* trajectories regardless
of absolute volume.

```{r indexed-comparison, fig.height=7}
indexed_trends <- crime_data %>%
  count(YEAR, TYPE) %>%
  group_by(TYPE) %>%
  mutate(
    base_val = first(n),
    index = round(n / base_val * 100, 0)
  ) %>%
  ungroup() %>%
  filter(TYPE %in% top_types)

base_year <- min(indexed_trends$YEAR)

ggplot(indexed_trends, aes(x = YEAR, y = index, color = TYPE, group = TYPE)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "grey60", linewidth = 0.4) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = unique(indexed_trends$YEAR)) +
  scale_color_brewer(palette = "Set2", name = "Crime Type") +
  labs(
    title = glue::glue("Indexed Crime Trends (Base Year {base_year} = 100)"),
    subtitle = "Values above 100 = growth vs base year; below 100 = decline",
    x = NULL, y = glue::glue("Index ({base_year} = 100)"),
    caption = "Top 8 crime types. Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right", panel.grid.major.x = element_line(color = "grey92"))
```

> **Part 2 takeaway:** Crime types are not changing uniformly. The chi-squared test
> above confirms whether the overall *mix* has shifted. Examining types individually
> reveals which categories drive the aggregate trend. The next section maps these
> patterns geographically.

---

# Part 3: Neighbourhood Analysis {.tabset .tabset-fade}

Geographic analysis reveals *where* crime concentrates and how neighbourhood-level
patterns differ from the aggregate trend. `r paste(EXCLUDE_HOODS, collapse = " and ")`
is excluded from these comparisons because its disproportionate volume compresses
the visual scale for other areas.

Click the tabs below to explore different neighbourhood perspectives.

## Top 10 (`r LATEST_YEAR`)

```{r top-neighbourhoods-latest}
top_hoods_latest <- crime_data %>%
  filter(YEAR == LATEST_YEAR, !NEIGHBOURHOOD %in% EXCLUDE_HOODS) %>%
  count(NEIGHBOURHOOD, sort = TRUE) %>%
  head(TOP_N_HOODS) %>%
  mutate(share = round(n / sum(n) * 100, 1))

ggplot(top_hoods_latest,
       aes(x = reorder(NEIGHBOURHOOD, n), y = n, fill = n)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(comma(n), " (", share, "%)")),
            hjust = -0.05, size = 3.3, fontface = "bold") +
  coord_flip() +
  scale_fill_viridis_c(option = "D", direction = -1) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.2))) +
  labs(
    title = glue::glue("Top {TOP_N_HOODS} Neighbourhoods by Crime Volume ({LATEST_YEAR})"),
    subtitle = glue::glue("{paste(EXCLUDE_HOODS, collapse = ', ')} excluded; labels show count and share of top {TOP_N_HOODS}"),
    x = NULL, y = "Total Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

## Year Comparison

```{r neighbourhood-facet, fig.height=10, fig.width=12}
neighbourhood_data <- crime_data %>%
  filter(YEAR %in% COMPARE_YEARS,
         !NEIGHBOURHOOD %in% EXCLUDE_HOODS) %>%
  count(YEAR, NEIGHBOURHOOD)

ggplot(neighbourhood_data,
       aes(x = reorder(NEIGHBOURHOOD, n), y = n, fill = n)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ YEAR, scales = "free_x", ncol = 3) +
  scale_fill_viridis_c(option = "D") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Total Crimes by Neighbourhood",
    subtitle = glue::glue("Side-by-side comparison: {paste(COMPARE_YEARS, collapse = ', ')} ({paste(EXCLUDE_HOODS, collapse = ', ')} excluded)"),
    x = NULL, y = "Total Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(axis.text.y = element_text(size = 7.5))
```

## Concentration

How concentrated is crime? The chart below shows cumulative crime share as we move
from the lowest-crime to the highest-crime neighbourhood. A steep curve at the right
indicates that a small number of neighbourhoods account for a large share of crime.

```{r crime-concentration, fig.height=6}
hood_concentration <- crime_data %>%
  filter(YEAR == LATEST_YEAR) %>%
  count(NEIGHBOURHOOD, sort = TRUE) %>%
  arrange(n) %>%
  mutate(
    cumulative_n = cumsum(n),
    cumulative_pct = cumulative_n / sum(n) * 100,
    hood_pct = row_number() / n() * 100
  )

# Find the 80/20 point
pct_80 <- hood_concentration %>%
  filter(cumulative_pct >= 80) %>%
  slice(1)
hoods_for_80 <- round(100 - pct_80$hood_pct, 0)

ggplot(hood_concentration, aes(x = hood_pct, y = cumulative_pct)) +
  geom_line(color = bc_accent, linewidth = 1.3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_hline(yintercept = 80, linetype = "dotted", color = bc_red, alpha = 0.6) +
  annotate("text", x = 20, y = 83,
           label = glue::glue("80% of crime from ~{hoods_for_80}% of neighbourhoods"),
           color = bc_red, size = 3.5, hjust = 0) +
  scale_x_continuous(labels = function(x) paste0(x, "%")) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = glue::glue("Crime Concentration Curve ({LATEST_YEAR})"),
    subtitle = "Cumulative % of crime vs cumulative % of neighbourhoods (Lorenz-style)",
    x = "Cumulative % of Neighbourhoods (low to high crime)",
    y = "Cumulative % of Total Crime",
    caption = "Dashed diagonal = perfect equality. Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(panel.grid.major.x = element_line(color = "grey92"))
```

## Change `r EARLIEST_YEAR`--`r LATEST_YEAR`

The overall crime trend (Part 1) showed the aggregate direction of change. But
this masks variation: some neighbourhoods may have improved even as others worsened.
The diverging bar chart below reveals which neighbourhoods drove the aggregate trend.

```{r neighbourhood-change, fig.height=8}
yr_early <- as.character(EARLIEST_YEAR)
yr_late  <- as.character(LATEST_YEAR)

hood_change <- crime_data %>%
  filter(YEAR %in% c(EARLIEST_YEAR, LATEST_YEAR),
         !NEIGHBOURHOOD %in% EXCLUDE_HOODS) %>%
  count(YEAR, NEIGHBOURHOOD) %>%
  pivot_wider(names_from = YEAR, values_from = n, values_fill = 0) %>%
  mutate(
    change_pct = safe_pct_change(.data[[yr_late]], .data[[yr_early]]),
    abs_change = .data[[yr_late]] - .data[[yr_early]],
    direction  = ifelse(change_pct > 0, "Increase", "Decrease")
  ) %>%
  filter(!is.na(change_pct)) %>%
  arrange(change_pct)

ggplot(hood_change,
       aes(x = reorder(NEIGHBOURHOOD, change_pct), y = change_pct, fill = direction)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.6, color = "grey40") +
  geom_text(
    aes(label = paste0(ifelse(change_pct > 0, "+", ""), change_pct, "%")),
    hjust = ifelse(hood_change$change_pct > 0, -0.1, 1.1),
    size = 3, fontface = "bold"
  ) +
  coord_flip() +
  scale_fill_manual(values = c("Increase" = bc_red, "Decrease" = bc_green)) +
  scale_y_continuous(labels = function(x) paste0(x, "%"),
                     expand = expansion(mult = 0.15)) +
  labs(
    title = glue::glue("Crime Change by Neighbourhood: {EARLIEST_YEAR} vs {LATEST_YEAR}"),
    subtitle = "Percentage change in total reported crimes; red = increase, green = decrease",
    x = NULL, y = "Change (%)",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r neighbourhood-change-table}
kable(
  hood_change %>%
    select(NEIGHBOURHOOD, all_of(c(yr_early, yr_late)), abs_change, change_pct, direction) %>%
    rename(
      Neighbourhood = NEIGHBOURHOOD,
      `Abs. Change` = abs_change,
      `Change (%)` = change_pct,
      Trend = direction
    ) %>%
    arrange(desc(`Change (%)`)),
  caption = glue::glue("Table 6: Crime volume change by neighbourhood, {EARLIEST_YEAR} to {LATEST_YEAR}"),
  format.args = list(big.mark = ",")
)
```

## Before & After

A dumbbell chart shows the absolute crime count in both `r EARLIEST_YEAR` and
`r LATEST_YEAR` for each neighbourhood, making it easy to see both the *level*
and the *direction* of change.

```{r dumbbell, fig.height=8}
dumbbell_data <- hood_change %>%
  arrange(desc(.data[[yr_late]])) %>%
  head(TOP_N_DUMBBELL)

ggplot(dumbbell_data) +
  geom_segment(aes(x = .data[[yr_early]], xend = .data[[yr_late]],
                   y = reorder(NEIGHBOURHOOD, .data[[yr_late]]),
                   yend = reorder(NEIGHBOURHOOD, .data[[yr_late]])),
               color = "grey70", linewidth = 0.8) +
  geom_point(aes(x = .data[[yr_early]], y = reorder(NEIGHBOURHOOD, .data[[yr_late]])),
             color = bc_accent, size = 3) +
  geom_point(aes(x = .data[[yr_late]], y = reorder(NEIGHBOURHOOD, .data[[yr_late]])),
             color = bc_red, size = 3) +
  scale_x_continuous(labels = comma) +
  labs(
    title = glue::glue("Neighbourhood Crime: {EARLIEST_YEAR} (blue) vs {LATEST_YEAR} (red)"),
    subtitle = glue::glue("Top {TOP_N_DUMBBELL} neighbourhoods by {LATEST_YEAR} volume; segment shows change"),
    x = "Total Reported Crimes", y = NULL,
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(panel.grid.major.x = element_line(color = "grey92"))
```

## Crime Profiles

Different neighbourhoods may have different *types* of crime, not just different
volumes. This heatmap shows what fraction of each neighbourhood's crime falls into
each category, revealing distinct neighbourhood "profiles."

```{r hood-profiles, fig.height=10, fig.width=12}
top_profile_hoods <- crime_data %>%
  filter(YEAR == LATEST_YEAR, !NEIGHBOURHOOD %in% EXCLUDE_HOODS) %>%
  count(NEIGHBOURHOOD, sort = TRUE) %>%
  head(TOP_N_HOODS_PROFILE) %>%
  pull(NEIGHBOURHOOD)

top_profile_types <- crime_data %>%
  filter(YEAR == LATEST_YEAR) %>%
  count(TYPE, sort = TRUE) %>%
  head(TOP_N_TYPES_PROFILE) %>%
  pull(TYPE)

hood_profile <- crime_data %>%
  filter(YEAR == LATEST_YEAR, NEIGHBOURHOOD %in% top_profile_hoods, TYPE %in% top_profile_types) %>%
  count(NEIGHBOURHOOD, TYPE) %>%
  group_by(NEIGHBOURHOOD) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  ungroup()

ggplot(hood_profile, aes(x = TYPE, y = NEIGHBOURHOOD, fill = pct)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(aes(label = paste0(pct, "%"),
                color = pct > median(pct)),
            size = 3.2, fontface = "bold", show.legend = FALSE) +
  scale_color_manual(values = c("TRUE" = "white", "FALSE" = "grey20")) +
  scale_fill_viridis_c(option = "D", name = "Share (%)") +
  labs(
    title = glue::glue("Neighbourhood Crime Profiles ({LATEST_YEAR})"),
    subtitle = "What % of each neighbourhood's crime falls into each type?",
    x = NULL, y = NULL,
    caption = glue::glue("Top {TOP_N_HOODS_PROFILE} neighbourhoods x top {TOP_N_TYPES_PROFILE} crime types. Source: BC municipal crime data")
  ) +
  theme_bc_crime() +
  theme(
    panel.grid = element_blank(),
    legend.position = "right",
    axis.text.x = element_text(angle = 35, hjust = 1, size = 9)
  )
```

## Neighbourhood Clusters

Which neighbourhoods have similar crime profiles? Hierarchical clustering groups
areas by the *composition* of their crime (not total volume), revealing natural
groupings that may correspond to urban planning zones or socioeconomic patterns.

```{r neighbourhood-clustering, fig.height=8, fig.width=10}
# Build a crime type proportion matrix per neighbourhood
cluster_data <- crime_data %>%
  filter(YEAR == LATEST_YEAR, !NEIGHBOURHOOD %in% EXCLUDE_HOODS) %>%
  count(NEIGHBOURHOOD, TYPE) %>%
  group_by(NEIGHBOURHOOD) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  select(NEIGHBOURHOOD, TYPE, pct) %>%
  pivot_wider(names_from = TYPE, values_from = pct, values_fill = 0)

# Need at least 3 neighbourhoods to cluster meaningfully
if (nrow(cluster_data) >= 3) {
  cluster_mat <- as.matrix(cluster_data[, -1])
  rownames(cluster_mat) <- cluster_data$NEIGHBOURHOOD

  # Hierarchical clustering with Ward's method
  dist_mat <- dist(cluster_mat, method = "euclidean")
  hc <- hclust(dist_mat, method = "ward.D2")

  # Cut into groups (use 3-5 clusters depending on data)
  n_clusters <- min(5, nrow(cluster_data) - 1)
  clusters <- cutree(hc, k = n_clusters)

  # Plot dendrogram
  plot(hc, hang = -1, cex = 0.7,
       main = glue::glue("Neighbourhood Crime Profile Clusters ({LATEST_YEAR})"),
       sub = "Ward's method on crime type proportions",
       xlab = "", ylab = "Distance")
  rect.hclust(hc, k = n_clusters, border = bc_palette[1:n_clusters])
}
```

```{r cluster-summary, results='asis'}
if (exists("clusters") && length(clusters) > 0) {
  cluster_df <- data.frame(
    NEIGHBOURHOOD = names(clusters),
    Cluster = clusters,
    stringsAsFactors = FALSE
  ) %>% arrange(Cluster, NEIGHBOURHOOD)

  cat("\n**Cluster memberships:**\n\n")
  for (cl in sort(unique(cluster_df$Cluster))) {
    members <- cluster_df %>% filter(Cluster == cl) %>% pull(NEIGHBOURHOOD)
    cat(glue::glue("- **Group {cl}** ({length(members)} areas): {paste(members, collapse = ', ')}\n"))
  }
  cat("\n")
  cat("> These clusters reflect similarity in crime *composition* (what proportion of each crime type), ")
  cat("not total volume. Two neighbourhoods in the same cluster have similar crime profiles.\n\n")
}
```

---

# Spotlight: Break-and-Enter Analysis

As a worked example of how to drill deeper into a specific crime category, this
section examines break-and-enter offences (both commercial and residential) in
more detail. This pattern can be replicated for any crime type of interest.

```{r be-spotlight, fig.height=6}
be_types <- c("Break and Enter Commercial", "Break and Enter Residential/Other")

be_data <- crime_data %>%
  filter(TYPE %in% be_types) %>%
  count(YEAR, TYPE)

ggplot(be_data, aes(x = YEAR, y = n, color = TYPE, group = TYPE)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_text(aes(label = comma(n)), vjust = -1, size = 3, show.legend = FALSE) +
  scale_x_continuous(breaks = unique(be_data$YEAR)) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0.05, 0.15))) +
  scale_color_manual(values = c("Break and Enter Commercial" = bc_accent,
                                 "Break and Enter Residential/Other" = bc_red),
                     name = NULL) +
  labs(
    title = "Break-and-Enter Trends: Commercial vs Residential",
    subtitle = "Tracking both sub-categories independently over time",
    x = NULL, y = "Number of Incidents",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "bottom")
```

```{r be-by-neighbourhood, fig.height=7}
top_be_hoods <- crime_data %>%
  filter(TYPE %in% be_types, YEAR == LATEST_YEAR,
         !NEIGHBOURHOOD %in% EXCLUDE_HOODS) %>%
  count(NEIGHBOURHOOD, sort = TRUE) %>%
  head(TOP_N_HOODS) %>%
  pull(NEIGHBOURHOOD)

be_by_hood <- crime_data %>%
  filter(TYPE %in% be_types, YEAR == LATEST_YEAR,
         NEIGHBOURHOOD %in% top_be_hoods) %>%
  count(NEIGHBOURHOOD, TYPE) %>%
  group_by(NEIGHBOURHOOD) %>%
  mutate(total = sum(n)) %>%
  ungroup()

ggplot(be_by_hood,
       aes(x = reorder(NEIGHBOURHOOD, total), y = n, fill = TYPE)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = c("Break and Enter Commercial" = bc_accent,
                                "Break and Enter Residential/Other" = bc_red),
                    name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(
    title = glue::glue("Break-and-Enter by Neighbourhood ({max(crime_data$YEAR)})"),
    subtitle = "Top 10 neighbourhoods; stacked by commercial vs residential",
    x = NULL, y = "Number of Incidents",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "bottom")
```

---

# Part 4: Seasonal and Monthly Patterns

Crime is not uniformly distributed throughout the year. Understanding monthly
patterns can inform resource allocation and public safety planning.

```{r monthly-check, include=FALSE}
# has_month is defined in the config chunk above
```

```{r monthly-analysis, eval=has_month, fig.height=7}
month_labels <- c("Jan","Feb","Mar","Apr","May","Jun",
                   "Jul","Aug","Sep","Oct","Nov","Dec")

monthly_crime <- crime_data %>%
  count(YEAR, MONTH) %>%
  mutate(MONTH_label = factor(month_labels[MONTH], levels = month_labels))

monthly_avg <- monthly_crime %>%
  group_by(MONTH, MONTH_label) %>%
  summarise(avg_n = mean(n), .groups = "drop")

ggplot(monthly_crime, aes(x = MONTH_label, y = n, group = YEAR, color = factor(YEAR))) +
  geom_line(alpha = 0.5, linewidth = 0.7) +
  geom_line(data = monthly_avg, aes(x = MONTH_label, y = avg_n, group = 1),
            color = "black", linewidth = 1.5, linetype = "solid", inherit.aes = FALSE) +
  scale_color_brewer(palette = "Set1", name = "Year") +
  labs(
    title = "Monthly Crime Patterns by Year",
    subtitle = "Thin coloured lines = individual years; thick black line = multi-year average",
    x = NULL, y = "Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right")
```

```{r monthly-heatmap, eval=has_month, fig.height=6}
ggplot(monthly_crime, aes(x = MONTH_label, y = factor(YEAR), fill = n)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = comma(n)), size = 3, color = "white", fontface = "bold") +
  scale_fill_viridis_c(option = "B", name = "Crimes", labels = comma) +
  labs(
    title = "Crime Volume Heatmap: Month x Year",
    subtitle = "Darker cells indicate higher crime volume",
    x = NULL, y = NULL,
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(panel.grid = element_blank(), legend.position = "right")
```

```{r no-month-note, eval=!has_month, results='asis'}
cat("*Monthly analysis is not available because the dataset does not contain a MONTH column.*\n\n")
```

---

# Part 5: Crime Severity Index

Not all crimes carry equal weight. A simple severity-weighted index gives a more
nuanced view than raw counts alone. The weights below are illustrative and based
on common criminal justice severity scales.

```{r severity-weights-table}
severity_weights <- data.frame(
  TYPE = c("Homicide", "Offence Against a Person",
           "Break and Enter Commercial", "Break and Enter Residential/Other",
           "Vehicle Collision or Pedestrian Struck (with Fatality)",
           "Vehicle Collision or Pedestrian Struck (with Injury)"),
  weight = c(10, 7, 5, 5, 9, 4),
  stringsAsFactors = FALSE
)

kable(
  severity_weights %>% arrange(desc(weight)) %>%
    rename(`Crime Type` = TYPE, `Severity Weight` = weight),
  caption = "Table 7: Severity weight assignments (all other types default to 2)"
)
```

```{r severity-index, fig.height=6}

crime_with_weight <- crime_data %>%
  left_join(severity_weights, by = "TYPE") %>%
  mutate(weight = ifelse(is.na(weight), 2, weight))

severity_by_year <- crime_with_weight %>%
  group_by(YEAR) %>%
  summarise(
    total_crimes = n(),
    severity_score = sum(weight),
    avg_severity = round(mean(weight), 2),
    .groups = "drop"
  )

ggplot(severity_by_year, aes(x = YEAR)) +
  geom_line(aes(y = avg_severity), color = bc_red, linewidth = 1.3) +
  geom_point(aes(y = avg_severity), color = bc_red, size = 3.5,
             shape = 21, fill = "white", stroke = 1.5) +
  geom_text(aes(y = avg_severity, label = avg_severity),
            vjust = -1.3, size = 3.5, color = "grey30") +
  scale_x_continuous(breaks = severity_by_year$YEAR) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  labs(
    title = "Average Crime Severity Index by Year",
    subtitle = "Weighted index: Homicide=10, Person offences=7, B&E=5, default=2",
    x = NULL, y = "Average Severity Score",
    caption = "Note: Weights are illustrative. Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r severity-table}
kable(
  severity_by_year %>%
    rename(Year = YEAR, `Total Crimes` = total_crimes,
           `Severity Sum` = severity_score, `Avg Severity` = avg_severity),
  caption = "Table 8: Annual severity index",
  format.args = list(big.mark = ",")
)
```

---

# Part 6: Trend Projections

Crime trends are not destiny, but understanding where current trajectories lead
can help inform resource planning. This section uses simple statistical models to
project crime volumes forward, with appropriate uncertainty bounds.

> **Important caveat:** These projections assume existing trends continue unchanged.
> Policy interventions, economic shifts, or demographic changes can alter trajectories
> substantially. Treat these as *baseline expectations*, not predictions.

```{r forecast-overall, fig.height=7, fig.alt="Fan chart showing crime volume projection with 80 percent prediction interval"}
# Simple exponential smoothing forecast using Holt's method (linear trend)
forecast_years <- params$forecast_years
future_years <- seq(LATEST_YEAR + 1, LATEST_YEAR + forecast_years)

# Fit Holt's linear trend model manually (no forecast package dependency)
annual_n <- crime_by_year$n
n_obs <- length(annual_n)

# Use linear regression as a robust baseline forecast
forecast_model <- lm(n ~ YEAR, data = crime_by_year)
forecast_vals <- predict(forecast_model,
                         newdata = data.frame(YEAR = future_years),
                         interval = "prediction", level = 0.80)

forecast_df <- data.frame(
  YEAR = future_years,
  n = forecast_vals[, "fit"],
  lower = forecast_vals[, "lwr"],
  upper = forecast_vals[, "upr"],
  type = "Projected"
)

# Naive forecast (carry forward last value)
naive_val <- tail(crime_by_year$n, 1)

# Combine historical + forecast
historical_df <- crime_by_year %>%
  select(YEAR, n) %>%
  mutate(lower = n, upper = n, type = "Historical")

combined_df <- bind_rows(historical_df, forecast_df)

ggplot(combined_df, aes(x = YEAR, y = n)) +
  geom_ribbon(data = forecast_df, aes(ymin = lower, ymax = upper),
              fill = bc_accent, alpha = 0.15) +
  geom_line(data = historical_df, color = bc_accent, linewidth = 1.2) +
  geom_point(data = historical_df, color = bc_accent, size = 3, shape = 21,
             fill = "white", stroke = 1.3) +
  geom_line(data = forecast_df, color = bc_accent, linewidth = 1.2, linetype = "dashed") +
  geom_point(data = forecast_df, color = bc_red, size = 3.5, shape = 17) +
  geom_hline(yintercept = naive_val, linetype = "dotted", color = "grey50") +
  annotate("text", x = LATEST_YEAR + forecast_years, y = naive_val,
           label = glue::glue("Naive baseline ({comma(naive_val)})"),
           hjust = 1, vjust = -0.8, size = 3, color = "grey50") +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0.05, 0.12))) +
  scale_x_continuous(breaks = c(crime_by_year$YEAR, future_years)) +
  labs(
    title = glue::glue("Crime Volume Projection: {LATEST_YEAR + 1}\u2013{LATEST_YEAR + forecast_years}"),
    subtitle = "Linear trend extrapolation with 80% prediction interval (shaded); dotted = naive baseline",
    x = NULL, y = "Projected Crimes",
    caption = "Projection: linear regression on historical data. Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r forecast-interpretation, results='asis'}
proj_latest <- round(forecast_vals[nrow(forecast_vals), "fit"], 0)
proj_lower  <- round(forecast_vals[nrow(forecast_vals), "lwr"], 0)
proj_upper  <- round(forecast_vals[nrow(forecast_vals), "upr"], 0)
proj_change <- safe_pct_change(proj_latest, naive_val)

cat(glue::glue(
  "> **Projection summary:** If the linear trend continues, {LATEST_YEAR + forecast_years} ",
  "crime volume is projected at approximately **{comma(proj_latest)}** incidents ",
  "(80% interval: {comma(proj_lower)}\u2013{comma(proj_upper)}). ",
  "This represents a **{format_change(proj_change)}** change relative to the ",
  "naive baseline of {comma(naive_val)} ({LATEST_YEAR} actuals carried forward).\n\n"
))
```

---

# Summary and Key Findings

```{r summary-stats, results='asis'}
total_latest  <- crime_data %>% filter(YEAR == max(YEAR)) %>% nrow()
total_earliest <- crime_data %>% filter(YEAR == min(YEAR)) %>% nrow()
overall_change <- safe_pct_change(total_latest, total_earliest)
latest_year    <- max(crime_data$YEAR)
earliest_year  <- min(crime_data$YEAR)

top_type <- crime_data %>%
  filter(YEAR == latest_year) %>%
  count(TYPE, sort = TRUE) %>%
  slice(1)

worst_hood <- hood_change %>% arrange(desc(change_pct)) %>% slice(1)
best_hood  <- hood_change %>% arrange(change_pct) %>% slice(1)

cat(glue::glue("
## Key Takeaways

1. **Overall trajectory:** Between {earliest_year} and {latest_year}, total reported
   crimes changed by **{ifelse(overall_change > 0, '+', '')}{overall_change}%**
   (from {comma(total_earliest)} to {comma(total_latest)} incidents).

2. **Dominant crime type ({latest_year}):** *{top_type$TYPE}* led with
   **{comma(top_type$n)}** incidents.

3. **Biggest neighbourhood increase:** *{worst_hood$NEIGHBOURHOOD}* saw a
   **+{worst_hood$change_pct}%** increase from {earliest_year} to {latest_year}.

4. **Biggest neighbourhood decrease:** *{best_hood$NEIGHBOURHOOD}* saw a
   **{best_hood$change_pct}%** decrease over the same period.

5. **Post-COVID context:** The 2020\u20132023 window captures pandemic-era disruptions
   and the subsequent recovery in reported crime volumes.

6. **Severity nuance:** Raw crime counts do not capture severity. The weighted
   severity index provides a complementary lens showing whether the *mix* of
   offences is becoming more or less serious over time.
"))
```

## What to Watch

```{r forward-looking, results='asis'}
# Identify the fastest-growing crime type
type_growth <- crime_data %>%
  filter(YEAR %in% c(earliest_year, latest_year)) %>%
  count(YEAR, TYPE) %>%
  pivot_wider(names_from = YEAR, values_from = n, values_fill = 0) %>%
  mutate(change = safe_pct_change(.data[[as.character(latest_year)]],
                                   .data[[as.character(earliest_year)]])) %>%
  filter(!is.na(change)) %>%
  arrange(desc(change))

fastest_growing <- type_growth %>% slice(1)
fastest_declining <- type_growth %>% arrange(change) %>% slice(1)

cat(glue::glue("
Based on the trends identified in this report, the following areas warrant
continued monitoring:

1. **Fastest-growing crime type:** *{fastest_growing$TYPE}*
   ({ifelse(fastest_growing$change > 0, '+', '')}{fastest_growing$change}% change,
   {earliest_year}\u2013{latest_year}). If this trend continues, it will become an
   increasingly large share of overall crime.

2. **Fastest-declining crime type:** *{fastest_declining$TYPE}*
   ({fastest_declining$change}% change). Worth investigating whether this reflects
   genuine reduction or reclassification.

3. **Neighbourhood hotspots:** The crime concentration analysis (Part 3) showed
   that ~{hoods_for_80}% of neighbourhoods account for 80% of crime. Targeted
   interventions in these areas could have outsized impact.

4. **Post-pandemic trajectory:** With {latest_year} being the latest data point,
   the next 1\u20132 years will reveal whether pandemic-era shifts are permanent or
   temporary.
"))
```

---

## Methodology Notes

### Data Source and Provenance

The analysis draws on publicly available municipal crime statistics from
British Columbia, Canada. Each record in the dataset represents one reported
criminal incident classified by year, offence type, and neighbourhood.

### Temporal Coverage

- **Period:** `r EARLIEST_YEAR` to `r LATEST_YEAR` (`r LATEST_YEAR - EARLIEST_YEAR + 1` years)
- **Granularity:** Annual aggregation for most analyses; monthly where MONTH
  column is available
- **Notable period:** 2020 captures COVID-19 pandemic-related disruptions

### Variable Definitions

| Variable | Definition |
|----------|------------|
| YEAR | Calendar year of the reported crime |
| TYPE | Offence category as classified by police reporting standards |
| NEIGHBOURHOOD | Geographic area (police reporting boundary) |
| MONTH | Calendar month (1--12), when available |

### Analytical Methods

| Method | Application | Section |
|--------|-------------|---------|
| Linear regression | Overall trend direction + significance | Part 1 |
| Chi-squared test | Crime type distribution change | Part 2 |
| Lorenz curve | Geographic concentration analysis | Part 3 |
| Hierarchical clustering (Ward's D2) | Neighbourhood crime profile grouping | Part 3 |
| Pearson correlation | Crime type co-movement | Part 2 |
| Severity-weighted index | Qualitative crime weighting | Part 5 |
| Linear extrapolation | Trend projection with prediction intervals | Part 6 |

### Exclusions and Assumptions

- `r paste(EXCLUDE_HOODS, collapse = " and ")` is excluded from neighbourhood
  comparisons due to disproportionate volume that compresses visual scales.
- Severity weights are **illustrative** and should be calibrated against
  official severity indices (e.g., Statistics Canada's Crime Severity Index)
  before policy application.
- All crime types not explicitly assigned a severity weight default to 2.

### Limitations

1. **Reported vs. actual crime:** This analysis uses *reported* crime counts.
   Under-reporting varies by crime type (e.g., sexual assault is historically
   under-reported) and can change over time due to policy or cultural shifts.
2. **No population normalization:** Neighbourhood-level population data was
   not available. Rising crime in a growing neighbourhood may reflect population
   growth rather than worsening safety.
3. **No causal claims:** This report identifies patterns and associations but
   does not establish causal relationships between interventions and outcomes.
4. **Temporal granularity:** Annual aggregation may mask within-year patterns;
   monthly analysis is provided where data permits.

### Reproducibility

```{r session-info, echo=FALSE, results='asis'}
cat(glue::glue(
  "- **R version:** {getRversion()}\n",
  "- **Platform:** {R.version$platform}\n",
  "- **Key packages:** ggplot2 {packageVersion('ggplot2')}, ",
  "dplyr {packageVersion('dplyr')}, ",
  "tidyr {packageVersion('tidyr')}, ",
  "scales {packageVersion('scales')}, ",
  "knitr {packageVersion('knitr')}\n",
  "- **Report generated:** {Sys.time()}\n"
))
```

---

# Appendix {.tabset}

## Glossary

| Term | Definition |
|------|------------|
| **Crime Severity Index (CSI)** | A Statistics Canada measure weighting crimes by their seriousness, based on average sentences |
| **Lorenz curve** | A graphical representation of inequality, showing the cumulative share of a quantity vs. cumulative share of the population |
| **Hierarchical clustering** | An unsupervised learning method that builds nested clusters by progressively merging or splitting groups |
| **Ward's method** | A clustering linkage criterion that minimizes total within-cluster variance at each merge step |
| **YoY (Year-over-Year)** | Comparison of a metric in one year to the same metric in the previous year |
| **Indexed comparison** | Expressing values relative to a base period (set to 100) to compare trends across different magnitudes |
| **IQR (Interquartile Range)** | The range between the 25th and 75th percentiles; used for outlier detection |
| **Chi-squared test** | A statistical test assessing whether the distribution of categorical data differs from expectation |
| **R-squared (R\u00b2)** | The proportion of variance in the outcome variable explained by the model |
| **Prediction interval** | A range within which a future observation is expected to fall with a given probability |

## Full Data Summary

```{r appendix-full-table}
full_summary <- crime_data %>%
  count(YEAR, TYPE) %>%
  pivot_wider(names_from = YEAR, values_from = n, values_fill = 0) %>%
  arrange(TYPE)

if (has_DT && knitr::is_html_output()) {
  DT::datatable(
    full_summary,
    caption = "Appendix A: Complete crime counts by type and year",
    options = list(pageLength = 15, scrollX = TRUE),
    rownames = FALSE,
    filter = "top"
  ) %>% DT::formatCurrency(names(full_summary)[-1], currency = "", digits = 0)
} else {
  kable(full_summary,
        caption = "Appendix A: Complete crime counts by type and year",
        format.args = list(big.mark = ","))
}
```

```{r appendix-hood-table}
hood_summary <- crime_data %>%
  count(YEAR, NEIGHBOURHOOD) %>%
  pivot_wider(names_from = YEAR, values_from = n, values_fill = 0) %>%
  arrange(NEIGHBOURHOOD)

if (has_DT && knitr::is_html_output()) {
  DT::datatable(
    hood_summary,
    caption = "Appendix B: Complete crime counts by neighbourhood and year",
    options = list(pageLength = 15, scrollX = TRUE),
    rownames = FALSE,
    filter = "top"
  ) %>% DT::formatCurrency(names(hood_summary)[-1], currency = "", digits = 0)
} else {
  kable(hood_summary,
        caption = "Appendix B: Complete crime counts by neighbourhood and year",
        format.args = list(big.mark = ","))
}
```

## References

1. Statistics Canada. *Police-reported crime statistics in Canada*. Annual release.
   [https://www150.statcan.gc.ca/n1/daily-quotidien/230727/dq230727b-eng.htm](https://www150.statcan.gc.ca/n1/daily-quotidien/230727/dq230727b-eng.htm)

2. BC Stats. *Population Estimates*. Government of British Columbia.
   [https://www2.gov.bc.ca/gov/content/data/statistics/people-population-community/population](https://www2.gov.bc.ca/gov/content/data/statistics/people-population-community/population)

3. Wickham, H. (2016). *ggplot2: Elegant Graphics for Data Analysis*. Springer-Verlag.
   [https://ggplot2.tidyverse.org](https://ggplot2.tidyverse.org)

4. R Core Team (`r R.version.string`). *R: A Language and Environment for Statistical Computing*.
   R Foundation for Statistical Computing, Vienna, Austria.
   [https://www.R-project.org/](https://www.R-project.org/)

---

```{r report-footer, echo=FALSE, results='asis'}
cat(glue::glue(
  "*Report generated on {Sys.Date()} using R {getRversion()}. ",
  "This analysis contains {exec_n_types} crime types across {exec_n_hoods} ",
  "neighbourhoods covering {comma(exec_total)} records ({exec_years[1]}\u2013{exec_years[2]}).*"
))
```
