---
title: "Exploring the Crime Landscape in British Columbia, Canada"
author: "Casper Kacper Ruta"
date: "2024-02-25"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
    fig_width: 10
    fig_height: 6
    fig_caption: true
    df_print: kable
  pdf_document:
    toc: true
    fig_width: 10
    fig_height: 6
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.retina = 2
)
```

# Introduction

This report examines crime trends across British Columbia, Canada, using publicly
available municipal crime data. The analysis covers the period from 2018 to 2023,
with a particular focus on post-COVID-19 trends (2020 onward) and neighbourhood-level
geographic patterns. The goal is to identify meaningful shifts in crime volume and
composition over time.

# Setup and Data Loading

```{r load-packages}
packages <- c("here", "readr", "stringr", "clock", "dplyr", "ggplot2",
              "rmarkdown", "knitr", "magrittr", "glue", "xfun", "fs",
              "tidyr", "scales")

installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(knitr)
```

```{r load-data}
crime_data <- read.csv("bc_crime.csv")
```

```{r custom-theme}
# Define a consistent theme used across all visualizations
theme_bc_crime <- function() {
  theme_minimal(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold", size = 16, hjust = 0),
      plot.subtitle = element_text(color = "grey40", size = 12, hjust = 0),
      plot.caption = element_text(color = "grey60", size = 9, hjust = 1),
      axis.title = element_text(face = "bold", size = 11),
      axis.text = element_text(size = 10),
      legend.position = "bottom",
      legend.title = element_text(face = "bold", size = 10),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      strip.text = element_text(face = "bold", size = 11),
      plot.margin = margin(10, 15, 10, 10)
    )
}

# Consistent colour palette
bc_blue_palette <- c("#08519c", "#3182bd", "#6baed6", "#9ecae1", "#c6dbef", "#deebf7")
```

# Data Overview

```{r data-summary}
cat("Dataset dimensions:", nrow(crime_data), "records x", ncol(crime_data), "columns\n")
cat("Years covered:", paste(sort(unique(crime_data$YEAR)), collapse = ", "), "\n")
cat("Crime types:", length(unique(crime_data$TYPE)), "\n")
cat("Neighbourhoods:", length(unique(crime_data$NEIGHBOURHOOD)), "\n")
```

```{r data-structure}
kable(
  head(crime_data, 10),
  caption = "Table 1: First 10 records of the BC crime dataset"
)
```

---

# Part 1: Overall Crime Trends

This section examines how total crime volume has changed year over year, providing
context for whether crime in British Columbia is increasing, decreasing, or holding
steady.

## Annual Crime Volume (2018--2023)

```{r annual-crime-trend}
crime_by_year <- crime_data %>%
  count(YEAR) %>%
  mutate(
    pct_change = round((n / lag(n) - 1) * 100, 1),
    label = comma(n)
  )

ggplot(crime_by_year, aes(x = YEAR, y = n)) +
  geom_line(color = "#08519c", linewidth = 1.2) +
  geom_point(color = "#08519c", size = 3) +
  geom_text(aes(label = label), vjust = -1.2, size = 3.5, color = "grey30") +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0.05, 0.12))) +
  scale_x_continuous(breaks = unique(crime_by_year$YEAR)) +
  labs(
    title = "Total Reported Crimes by Year",
    subtitle = "British Columbia, 2018\u20132023",
    x = "Year",
    y = "Number of Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r annual-change-table}
kable(
  crime_by_year %>% select(YEAR, n, pct_change) %>%
    rename(Year = YEAR, `Total Crimes` = n, `Year-over-Year Change (%)` = pct_change),
  caption = "Table 2: Annual crime counts with year-over-year percentage change",
  format.args = list(big.mark = ",")
)
```

## Year-over-Year Change

```{r yoy-bar}
crime_by_year_change <- crime_by_year %>% filter(!is.na(pct_change))

ggplot(crime_by_year_change, aes(x = factor(YEAR), y = pct_change,
                                  fill = pct_change > 0)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.5) +
  geom_text(aes(label = paste0(ifelse(pct_change > 0, "+", ""), pct_change, "%")),
            vjust = ifelse(crime_by_year_change$pct_change > 0, -0.5, 1.5),
            size = 3.8, fontface = "bold") +
  scale_fill_manual(values = c("TRUE" = "#c0392b", "FALSE" = "#27ae60")) +
  scale_y_continuous(labels = function(x) paste0(x, "%"),
                     expand = expansion(mult = c(0.15, 0.15))) +
  labs(
    title = "Year-over-Year Change in Total Crimes",
    subtitle = "Red = increase, Green = decrease",
    x = "Year",
    y = "Change from Previous Year (%)",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

---

# Part 2: Crime Type Analysis

Understanding the composition of crime by type reveals which categories drive
overall trends and whether certain crime types are becoming more or less prevalent.

## Crime Counts by Type and Year

```{r crime-type-summary}
crime_counts <- crime_data %>%
  filter(YEAR >= 2019) %>%
  group_by(YEAR, TYPE) %>%
  summarise(Count = n(), .groups = "drop") %>%
  arrange(YEAR, desc(Count))

crime_type_wide <- crime_counts %>%
  pivot_wider(names_from = YEAR, values_from = Count, values_fill = 0)

kable(
  crime_type_wide,
  caption = "Table 3: Crime counts by type and year (2019\u20132023)",
  format.args = list(big.mark = ",")
)
```

## Crime Type Distribution (Stacked Area)

```{r crime-type-area, fig.height=7}
top_types <- crime_data %>%
  count(TYPE, sort = TRUE) %>%
  head(8) %>%
  pull(TYPE)

crime_type_trend <- crime_data %>%
  mutate(TYPE_grouped = ifelse(TYPE %in% top_types, TYPE, "Other")) %>%
  count(YEAR, TYPE_grouped) %>%
  rename(TYPE = TYPE_grouped)

ggplot(crime_type_trend, aes(x = YEAR, y = n, fill = TYPE)) +
  geom_area(alpha = 0.8) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = unique(crime_type_trend$YEAR)) +
  labs(
    title = "Crime Composition Over Time",
    subtitle = "Top 8 crime types shown individually; remaining types grouped as 'Other'",
    x = "Year",
    y = "Number of Crimes",
    fill = "Crime Type",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right")
```

## Homicide Trends (Post-COVID Focus)

Homicides, while a small fraction of total crime, are the most severe category
and deserve dedicated attention.

```{r homicide-trend}
homicide_count <- crime_data %>%
  filter(TYPE == "Homicide") %>%
  count(YEAR) %>%
  mutate(label = n)

ggplot(homicide_count, aes(x = YEAR, y = n)) +
  geom_col(fill = "#c0392b", width = 0.6, alpha = 0.85) +
  geom_text(aes(label = n), vjust = -0.5, fontface = "bold", size = 4) +
  scale_x_continuous(breaks = unique(homicide_count$YEAR)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Homicide Count by Year",
    subtitle = "British Columbia \u2014 all neighbourhoods",
    x = "Year",
    y = "Number of Homicides",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r homicide-table}
kable(
  homicide_count %>% select(YEAR, n) %>% rename(Year = YEAR, Homicides = n),
  caption = "Table 4: Annual homicide counts"
)
```

---

# Part 3: Neighbourhood Analysis

This section explores geographic variation in crime across neighbourhoods, helping
identify which areas experience disproportionate crime levels and how those patterns
shift over time.

## Neighbourhood Crime Comparison (Faceted by Year)

Rather than examining each year in isolation, the faceted view below allows direct
visual comparison across years.

```{r neighbourhood-facet, fig.height=10, fig.width=12}
neighbourhood_data <- crime_data %>%
  filter(YEAR %in% c(2018, 2019, 2023),
         NEIGHBOURHOOD != "Central Business District") %>%
  count(YEAR, NEIGHBOURHOOD) %>%
  group_by(YEAR) %>%
  mutate(rank = dense_rank(desc(n))) %>%
  ungroup()

ggplot(neighbourhood_data,
       aes(x = reorder(NEIGHBOURHOOD, n), y = n, fill = n)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ YEAR, scales = "free_x", ncol = 3) +
  scale_fill_viridis_c(option = "D") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Total Crimes by Neighbourhood",
    subtitle = "Comparing 2018, 2019, and 2023 (Central Business District excluded)",
    x = NULL,
    y = "Total Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(
    axis.text.y = element_text(size = 8),
    strip.text = element_text(size = 13)
  )
```

## Top 10 Neighbourhoods by Crime Volume (2023)

```{r top-neighbourhoods-2023}
top_neighbourhoods_2023 <- crime_data %>%
  filter(YEAR == 2023, NEIGHBOURHOOD != "Central Business District") %>%
  count(NEIGHBOURHOOD, sort = TRUE) %>%
  head(10)

ggplot(top_neighbourhoods_2023,
       aes(x = reorder(NEIGHBOURHOOD, n), y = n, fill = n)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = comma(n)), hjust = -0.1, size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_viridis_c(option = "D", direction = -1) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top 10 Neighbourhoods by Crime Volume (2023)",
    subtitle = "Central Business District excluded",
    x = NULL,
    y = "Total Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

## Neighbourhood Trend: Change from 2018 to 2023

```{r neighbourhood-change, fig.height=8}
neighbourhood_change <- crime_data %>%
  filter(YEAR %in% c(2018, 2023),
         NEIGHBOURHOOD != "Central Business District") %>%
  count(YEAR, NEIGHBOURHOOD) %>%
  pivot_wider(names_from = YEAR, values_from = n, values_fill = 0) %>%
  mutate(
    change_pct = round((`2023` / `2018` - 1) * 100, 1),
    direction = ifelse(change_pct > 0, "Increase", "Decrease")
  ) %>%
  arrange(change_pct)

ggplot(neighbourhood_change,
       aes(x = reorder(NEIGHBOURHOOD, change_pct), y = change_pct, fill = direction)) +
  geom_col(width = 0.7, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.5) +
  coord_flip() +
  scale_fill_manual(values = c("Increase" = "#c0392b", "Decrease" = "#27ae60")) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Crime Change by Neighbourhood: 2018 vs 2023",
    subtitle = "Percentage change in total reported crimes",
    x = NULL,
    y = "Change (%)",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r neighbourhood-change-table}
kable(
  neighbourhood_change %>%
    select(NEIGHBOURHOOD, `2018`, `2023`, change_pct, direction) %>%
    rename(Neighbourhood = NEIGHBOURHOOD, `Change (%)` = change_pct, Trend = direction) %>%
    arrange(desc(`Change (%)`)),
  caption = "Table 5: Crime volume change by neighbourhood, 2018 to 2023",
  format.args = list(big.mark = ",")
)
```

---

# Part 4: Crime Type Heatmap

```{r crime-heatmap, fig.height=8, fig.width=10}
crime_type_year <- crime_data %>%
  count(YEAR, TYPE) %>%
  group_by(YEAR) %>%
  mutate(proportion = n / sum(n) * 100) %>%
  ungroup()

ggplot(crime_type_year, aes(x = factor(YEAR), y = reorder(TYPE, n), fill = proportion)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = round(proportion, 1)), size = 3, color = "white") +
  scale_fill_viridis_c(option = "C", name = "Share (%)") +
  labs(
    title = "Crime Type Proportions by Year",
    subtitle = "Cell values show the percentage share of each crime type within that year",
    x = "Year",
    y = NULL,
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(
    legend.position = "right",
    panel.grid = element_blank()
  )
```

---

# Summary and Key Findings

```{r summary-stats, results='asis'}
total_latest <- crime_data %>% filter(YEAR == max(YEAR)) %>% nrow()
total_earliest <- crime_data %>% filter(YEAR == min(YEAR)) %>% nrow()
overall_change <- round((total_latest / total_earliest - 1) * 100, 1)
latest_year <- max(crime_data$YEAR)
earliest_year <- min(crime_data$YEAR)

top_type <- crime_data %>%
  filter(YEAR == latest_year) %>%
  count(TYPE, sort = TRUE) %>%
  slice(1)

cat(glue::glue("
**Key findings from the analysis:**

1. **Overall trend:** From {earliest_year} to {latest_year}, total reported crimes
   changed by **{ifelse(overall_change > 0, '+', '')}{overall_change}%**
   ({comma(total_earliest)} to {comma(total_latest)} incidents).

2. **Most common crime type ({latest_year}):** {top_type$TYPE}
   with **{comma(top_type$n)}** reported incidents.

3. **Geographic concentration:** The Central Business District was excluded from
   neighbourhood comparisons due to its disproportionately high volume, which
   compressed the visual scale for other areas.

4. **Post-COVID patterns:** The 2020\u20132023 period shows how crime patterns shifted
   following the COVID-19 pandemic and subsequent recovery.
"))
```

---

*Report generated on `r Sys.Date()` using R `r getRversion()`.*
