---
title: "Exploring the Crime Landscape in British Columbia, Canada"
author: "Casper Kacper Ruta"
date: "2024-02-25"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    theme: flatly
    highlight: tango
    code_folding: hide
    fig_width: 10
    fig_height: 6
    fig_caption: true
    df_print: kable
    css: custom.css
  pdf_document:
    toc: true
    fig_width: 10
    fig_height: 6
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.retina = 2,
  out.width = "100%"
)
```

# Executive Summary {.tabset}

```{r exec-summary, echo=FALSE, results='asis'}
# Pre-compute key stats for the executive summary
exec_total   <- nrow(crime_data)
exec_years   <- range(crime_data$YEAR)
exec_latest  <- crime_data %>% filter(YEAR == max(YEAR)) %>% nrow()
exec_first   <- crime_data %>% filter(YEAR == min(YEAR)) %>% nrow()
exec_change  <- safe_pct_change(exec_latest, exec_first)
exec_top     <- crime_data %>% filter(YEAR == max(YEAR)) %>% count(TYPE, sort = TRUE) %>% slice(1)
exec_n_types <- length(unique(crime_data$TYPE))
exec_n_hoods <- length(unique(crime_data$NEIGHBOURHOOD))

cat(glue::glue("
| Metric | Value |
|:-------|------:|
| **Total records** | {comma(exec_total)} |
| **Period covered** | {exec_years[1]}\u2013{exec_years[2]} |
| **Crime types** | {exec_n_types} |
| **Neighbourhoods** | {exec_n_hoods} |
| **Latest year volume** | {comma(exec_latest)} |
| **Overall change** | {ifelse(exec_change > 0, '+', '')}{exec_change}% |
| **Top crime type ({exec_years[2]})** | {exec_top$TYPE} ({comma(exec_top$n)}) |
"))
```

---

# Introduction

This report examines crime trends across British Columbia, Canada, using publicly
available municipal crime data. The analysis spans **`r exec_years[1]`** to
**`r exec_years[2]`** and is organized into five parts:

1. **Overall Trends** --- aggregate crime volume and year-over-year momentum
2. **Crime Type Analysis** --- compositional breakdown by offence category
3. **Neighbourhood Analysis** --- geographic variation and hotspot identification
4. **Seasonal & Monthly Patterns** --- temporal rhythms within each year
5. **Crime Severity Index** --- a weighted view beyond raw counts

The overarching question: *How has crime in British Columbia evolved in the years
surrounding the COVID-19 pandemic, and which communities and crime types are driving
those changes?*

# Setup and Data Loading

```{r load-packages}
packages <- c("here", "readr", "stringr", "clock", "dplyr", "ggplot2",
              "rmarkdown", "knitr", "magrittr", "glue", "xfun", "fs",
              "tidyr", "scales")

installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(knitr)
```

```{r load-data}
crime_data <- read.csv("bc_crime.csv")
```

```{r custom-theme}
# ------------------------------------------------------------------
# Unified visual identity for all charts
# ------------------------------------------------------------------
theme_bc_crime <- function(base_size = 13) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      plot.title       = element_text(face = "bold", size = rel(1.25), hjust = 0,
                                      margin = margin(b = 4)),
      plot.subtitle    = element_text(color = "grey40", size = rel(0.9), hjust = 0,
                                      margin = margin(b = 10)),
      plot.caption     = element_text(color = "grey60", size = rel(0.7), hjust = 1,
                                      margin = margin(t = 8)),
      axis.title       = element_text(face = "bold", size = rel(0.85)),
      axis.text        = element_text(size = rel(0.8)),
      legend.position  = "bottom",
      legend.title     = element_text(face = "bold", size = rel(0.8)),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      strip.text       = element_text(face = "bold", size = rel(0.9)),
      strip.background = element_rect(fill = "grey95", color = NA),
      plot.margin      = margin(10, 15, 10, 10)
    )
}

# Palettes
bc_accent   <- "#08519c"
bc_red      <- "#c0392b"
bc_green    <- "#27ae60"
bc_palette  <- c("#08519c", "#3182bd", "#6baed6", "#9ecae1", "#c6dbef", "#deebf7")

# Helper: safe division avoiding Inf / NaN
safe_pct_change <- function(new, old) {
  ifelse(old == 0, NA_real_, round((new / old - 1) * 100, 1))
}

# Reusable horizontal bar chart for neighbourhood data
plot_hood_bar <- function(data, title, subtitle = NULL, show_labels = TRUE) {
  p <- ggplot(data, aes(x = reorder(NEIGHBOURHOOD, n), y = n, fill = n)) +
    geom_col(show.legend = FALSE) +
    coord_flip() +
    scale_fill_viridis_c(option = "D", direction = -1) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
    labs(title = title, subtitle = subtitle, x = NULL, y = "Total Reported Crimes",
         caption = "Source: BC municipal crime data") +
    theme_bc_crime()
  if (show_labels) {
    p <- p + geom_text(aes(label = comma(n)), hjust = -0.05, size = 3.2, fontface = "bold")
  }
  p
}
```

# Data Quality Check

Before analysis, it is good practice to validate the dataset for completeness
and consistency. This section runs automated checks.

```{r data-quality, results='asis'}
# Automated data quality checks
checks <- list()

# Check 1: Missing values
na_counts <- colSums(is.na(crime_data))
na_total  <- sum(na_counts)
checks$missing <- ifelse(na_total == 0,
  "**PASS** --- No missing values detected.",
  glue::glue("**WARNING** --- {na_total} missing values found: {paste(names(na_counts[na_counts > 0]), collapse = ', ')}"))

# Check 2: Year continuity
years <- sort(unique(crime_data$YEAR))
year_gaps <- setdiff(seq(min(years), max(years)), years)
checks$year_gaps <- ifelse(length(year_gaps) == 0,
  glue::glue("**PASS** --- Continuous year coverage ({min(years)}\u2013{max(years)})."),
  glue::glue("**WARNING** --- Missing years: {paste(year_gaps, collapse = ', ')}"))

# Check 3: Record count per year (check for suspiciously low years)
year_counts <- crime_data %>% count(YEAR)
min_year <- year_counts %>% filter(n == min(n))
max_year <- year_counts %>% filter(n == max(n))
ratio <- round(max_year$n[1] / min_year$n[1], 1)
checks$balance <- glue::glue(
  "**INFO** --- Year with fewest records: {min_year$YEAR[1]} ({comma(min_year$n[1])}). ",
  "Year with most: {max_year$YEAR[1]} ({comma(max_year$n[1])}). Ratio: {ratio}x.")

cat("| Check | Result |\n|:------|:-------|\n")
cat(glue::glue("| Missing values | {checks$missing} |\n"))
cat(glue::glue("| Year continuity | {checks$year_gaps} |\n"))
cat(glue::glue("| Record balance | {checks$balance} |\n"))
```

---

# Data Overview

```{r data-overview, results='asis'}
n_records <- format(nrow(crime_data), big.mark = ",")
n_types   <- length(unique(crime_data$TYPE))
n_hoods   <- length(unique(crime_data$NEIGHBOURHOOD))
yr_range  <- range(crime_data$YEAR)

cat(glue::glue(
  "The dataset contains **{n_records}** individual crime records spanning ",
  "**{yr_range[1]}** to **{yr_range[2]}**, encompassing **{n_types}** offence ",
  "categories across **{n_hoods}** neighbourhoods.\n\n"
))
```

```{r data-preview}
kable(
  head(crime_data, 8),
  caption = "Table 1: First 8 records of the BC crime dataset"
)
```

```{r data-column-summary}
col_summary <- data.frame(
  Column = names(crime_data),
  Class  = sapply(crime_data, class),
  Unique = sapply(crime_data, function(x) length(unique(x))),
  Missing = sapply(crime_data, function(x) sum(is.na(x))),
  row.names = NULL
)
kable(col_summary, caption = "Table 2: Column-level summary statistics")
```

---

# Part 1: Overall Crime Trends

This section tracks aggregate crime volume over time. Understanding the macro trend
is essential before drilling into specific crime types or neighbourhoods.

## Annual Crime Volume

```{r annual-crime-trend}
crime_by_year <- crime_data %>%
  count(YEAR) %>%
  mutate(
    pct_change = safe_pct_change(n, lag(n)),
    label = comma(n),
    cumulative = cumsum(n)
  )

ggplot(crime_by_year, aes(x = YEAR, y = n)) +
  annotate("rect", xmin = 2019.5, xmax = 2020.5,
           ymin = -Inf, ymax = Inf, fill = bc_red, alpha = 0.06) +
  annotate("text", x = 2020, y = max(crime_by_year$n) * 1.08,
           label = "COVID-19\nonset", color = bc_red, size = 3, fontface = "italic") +
  geom_ribbon(aes(ymin = min(n) * 0.95, ymax = n), fill = bc_accent, alpha = 0.08) +
  geom_line(color = bc_accent, linewidth = 1.3) +
  geom_point(color = bc_accent, size = 3.5, shape = 21, fill = "white", stroke = 1.5) +
  geom_text(aes(label = label), vjust = -1.4, size = 3.5, color = "grey25") +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0.03, 0.18))) +
  scale_x_continuous(breaks = crime_by_year$YEAR) +
  labs(
    title = "Total Reported Crimes by Year",
    subtitle = glue::glue("British Columbia, {yr_range[1]}\u2013{yr_range[2]} | Shaded band marks the COVID-19 pandemic year"),
    x = NULL, y = "Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r annual-change-table}
kable(
  crime_by_year %>%
    select(YEAR, n, pct_change) %>%
    rename(Year = YEAR, `Total Crimes` = n, `YoY Change (%)` = pct_change),
  caption = "Table 3: Annual crime counts with year-over-year percentage change",
  format.args = list(big.mark = ",")
)
```

## Year-over-Year Momentum

The bar chart below isolates the *direction and magnitude* of change each year.
Green bars indicate a year where crime fell relative to the prior year; red bars
indicate an increase.

```{r yoy-bar}
yoy_data <- crime_by_year %>% filter(!is.na(pct_change))

ggplot(yoy_data, aes(x = factor(YEAR), y = pct_change, fill = pct_change > 0)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.6, color = "grey40") +
  geom_text(
    aes(label = paste0(ifelse(pct_change > 0, "+", ""), pct_change, "%")),
    vjust = ifelse(yoy_data$pct_change > 0, -0.6, 1.6),
    size = 4, fontface = "bold"
  ) +
  scale_fill_manual(values = c("TRUE" = bc_red, "FALSE" = bc_green)) +
  scale_y_continuous(labels = function(x) paste0(x, "%"),
                     expand = expansion(mult = 0.2)) +
  labs(
    title = "Year-over-Year Change in Total Reported Crimes",
    subtitle = "Red = increase over prior year | Green = decrease",
    x = NULL, y = "Change (%)",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

---

# Part 2: Crime Type Analysis

Not all crimes are equal in severity or frequency. This section decomposes the
aggregate trend by offence type to understand *what kinds* of crime are rising or
falling.

## Crime Counts by Type and Year

```{r crime-type-table}
crime_counts <- crime_data %>%
  group_by(YEAR, TYPE) %>%
  summarise(Count = n(), .groups = "drop")

crime_type_wide <- crime_counts %>%
  pivot_wider(names_from = YEAR, values_from = Count, values_fill = 0)

kable(
  crime_type_wide,
  caption = glue::glue("Table 4: Crime counts by type and year ({yr_range[1]}\u2013{yr_range[2]})"),
  format.args = list(big.mark = ",")
)
```

## Crime Type Distribution (Latest Year)

Before examining trends, here is the overall breakdown for the most recent year.
This proportional bar chart replaces a pie chart --- it is easier to compare
relative sizes accurately.

```{r crime-type-proportion}
type_dist <- crime_data %>%
  filter(YEAR == max(YEAR)) %>%
  count(TYPE, sort = TRUE) %>%
  mutate(
    pct = round(n / sum(n) * 100, 1),
    label = ifelse(pct >= 2, paste0(TYPE, "\n", pct, "%"), "")
  )

ggplot(type_dist, aes(x = 1, y = n, fill = reorder(TYPE, -n))) +
  geom_col(width = 1, color = "white", linewidth = 0.3, show.legend = TRUE) +
  coord_flip() +
  scale_fill_brewer(palette = "Set3", name = "Crime Type") +
  labs(
    title = glue::glue("Crime Type Distribution ({max(crime_data$YEAR)})"),
    subtitle = "Proportional bar; each segment represents one crime type's share",
    x = NULL, y = NULL,
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.position = "right"
  )
```

## Crime Composition Over Time

The stacked area chart reveals how the *mix* of crime types shifts across years.
The top 8 categories are shown individually; less common types are grouped.

```{r crime-type-area, fig.height=7}
top_types <- crime_data %>%
  count(TYPE, sort = TRUE) %>%
  head(8) %>%
  pull(TYPE)

crime_type_trend <- crime_data %>%
  mutate(TYPE_grouped = ifelse(TYPE %in% top_types, TYPE, "Other")) %>%
  count(YEAR, TYPE_grouped) %>%
  rename(TYPE = TYPE_grouped)

ggplot(crime_type_trend, aes(x = YEAR, y = n, fill = TYPE)) +
  geom_area(alpha = 0.85, color = "white", linewidth = 0.3) +
  scale_fill_brewer(palette = "Set2", name = "Crime Type") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = unique(crime_type_trend$YEAR)) +
  labs(
    title = "Crime Composition Over Time",
    subtitle = "Top 8 crime types shown individually; remaining grouped as 'Other'",
    x = NULL, y = "Number of Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right")
```

## Crime Type Proportions Heatmap

This heatmap normalizes each year's data to show the *percentage share* of each
crime type, making it easy to spot shifts in crime composition independent of
overall volume changes.

```{r crime-heatmap, fig.height=9, fig.width=10}
crime_type_year <- crime_data %>%
  count(YEAR, TYPE) %>%
  group_by(YEAR) %>%
  mutate(proportion = round(n / sum(n) * 100, 1)) %>%
  ungroup()

# Order types by total volume for better readability
type_order <- crime_type_year %>%
  group_by(TYPE) %>%
  summarise(total = sum(n)) %>%
  arrange(total) %>%
  pull(TYPE)

crime_type_year <- crime_type_year %>%
  mutate(TYPE = factor(TYPE, levels = type_order))

ggplot(crime_type_year, aes(x = factor(YEAR), y = TYPE, fill = proportion)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(aes(label = ifelse(proportion >= 0.5, paste0(proportion, "%"), ""),
                color = proportion > median(proportion)),
            size = 3, fontface = "bold", show.legend = FALSE) +
  scale_color_manual(values = c("TRUE" = "white", "FALSE" = "grey20")) +
  scale_fill_viridis_c(option = "C", name = "Share of\nYear's Total (%)") +
  labs(
    title = "Crime Type Proportions by Year",
    subtitle = "Each cell shows the percentage share within that year; types ordered by total volume",
    x = "Year", y = NULL,
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right", panel.grid = element_blank())
```

## Homicide Trends

Homicides, while a small fraction of total crime, represent the most severe
category. The dedicated chart below tracks their year-over-year trajectory.

```{r homicide-trend}
homicide_count <- crime_data %>%
  filter(TYPE == "Homicide") %>%
  count(YEAR) %>%
  mutate(
    pct_change = safe_pct_change(n, lag(n)),
    avg = mean(n)
  )

ggplot(homicide_count, aes(x = YEAR, y = n)) +
  geom_hline(aes(yintercept = avg), linetype = "dashed", color = "grey50", linewidth = 0.5) +
  annotate("text", x = min(homicide_count$YEAR) + 0.3, y = homicide_count$avg[1] + 0.5,
           label = paste0("Average: ", round(homicide_count$avg[1], 1)),
           color = "grey50", size = 3.5, hjust = 0) +
  geom_col(fill = bc_red, width = 0.55, alpha = 0.85) +
  geom_text(aes(label = n), vjust = -0.6, fontface = "bold", size = 4.2) +
  scale_x_continuous(breaks = homicide_count$YEAR) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(
    title = "Homicide Count by Year",
    subtitle = "Dashed line = period average across all years",
    x = NULL, y = "Number of Homicides",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r homicide-table}
kable(
  homicide_count %>%
    select(YEAR, n, pct_change) %>%
    rename(Year = YEAR, Homicides = n, `YoY Change (%)` = pct_change),
  caption = "Table 5: Annual homicide counts with year-over-year change"
)
```

## Top Crime Types: Rank Change Over Time

This bump chart shows how the relative ranking of crime types has shifted. A rising
line means a crime type is becoming relatively more common.

```{r rank-bump, fig.height=7}
type_ranks <- crime_data %>%
  count(YEAR, TYPE) %>%
  group_by(YEAR) %>%
  mutate(rank = dense_rank(desc(n))) %>%
  ungroup() %>%
  filter(TYPE %in% top_types)

ggplot(type_ranks, aes(x = YEAR, y = rank, color = TYPE, group = TYPE)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2.5) +
  scale_y_reverse(breaks = 1:max(type_ranks$rank)) +
  scale_x_continuous(breaks = unique(type_ranks$YEAR)) +
  scale_color_brewer(palette = "Set2", name = "Crime Type") +
  labs(
    title = "Crime Type Rankings Over Time",
    subtitle = "Top 8 crime types by volume; rank 1 = most common that year",
    x = NULL, y = "Rank (1 = most common)",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right", panel.grid.major.x = element_line(color = "grey90"))
```

## Individual Crime Type Trends (Small Multiples)

Each panel below tracks one crime type independently, making it easy to see which
types are rising, falling, or volatile year to year.

```{r small-multiples, fig.height=10, fig.width=12}
type_trends <- crime_data %>%
  count(YEAR, TYPE) %>%
  group_by(TYPE) %>%
  mutate(
    first_val = first(n),
    indexed = round(n / first_val * 100, 0)
  ) %>%
  ungroup()

ggplot(type_trends, aes(x = YEAR, y = n)) +
  geom_line(color = bc_accent, linewidth = 0.8) +
  geom_point(color = bc_accent, size = 1.5) +
  facet_wrap(~ TYPE, scales = "free_y", ncol = 4) +
  scale_x_continuous(breaks = c(min(type_trends$YEAR), max(type_trends$YEAR))) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Individual Crime Type Trends",
    subtitle = "Each panel uses its own y-axis scale to highlight per-type patterns",
    x = NULL, y = NULL,
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime(base_size = 10) +
  theme(
    strip.text = element_text(size = 8),
    panel.grid.major.x = element_line(color = "grey92")
  )
```

## Indexed Comparison (Base Year = 100)

Comparing crime types of very different magnitudes (e.g., thousands of thefts vs
dozens of homicides) on the same scale is misleading. By indexing each type to
100 in the first available year, we can compare *relative* trajectories regardless
of absolute volume.

```{r indexed-comparison, fig.height=7}
indexed_trends <- crime_data %>%
  count(YEAR, TYPE) %>%
  group_by(TYPE) %>%
  mutate(
    base_val = first(n),
    index = round(n / base_val * 100, 0)
  ) %>%
  ungroup() %>%
  filter(TYPE %in% top_types)

base_year <- min(indexed_trends$YEAR)

ggplot(indexed_trends, aes(x = YEAR, y = index, color = TYPE, group = TYPE)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "grey60", linewidth = 0.4) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = unique(indexed_trends$YEAR)) +
  scale_color_brewer(palette = "Set2", name = "Crime Type") +
  labs(
    title = glue::glue("Indexed Crime Trends (Base Year {base_year} = 100)"),
    subtitle = "Values above 100 = growth vs base year; below 100 = decline",
    x = NULL, y = glue::glue("Index ({base_year} = 100)"),
    caption = "Top 8 crime types. Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right", panel.grid.major.x = element_line(color = "grey92"))
```

---

# Part 3: Neighbourhood Analysis

Geographic analysis reveals *where* crime concentrates and how neighbourhood-level
patterns differ from the aggregate trend. The Central Business District is excluded
from these comparisons because its disproportionate volume compresses the visual
scale for other areas.

## Top 10 Neighbourhoods (2023)

```{r top-neighbourhoods-2023}
top_hoods_2023 <- crime_data %>%
  filter(YEAR == 2023, NEIGHBOURHOOD != "Central Business District") %>%
  count(NEIGHBOURHOOD, sort = TRUE) %>%
  head(10) %>%
  mutate(share = round(n / sum(n) * 100, 1))

ggplot(top_hoods_2023,
       aes(x = reorder(NEIGHBOURHOOD, n), y = n, fill = n)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(comma(n), " (", share, "%)")),
            hjust = -0.05, size = 3.3, fontface = "bold") +
  coord_flip() +
  scale_fill_viridis_c(option = "D", direction = -1) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.2))) +
  labs(
    title = "Top 10 Neighbourhoods by Crime Volume (2023)",
    subtitle = "Central Business District excluded; labels show count and share of top 10",
    x = NULL, y = "Total Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

## Neighbourhood Comparison Across Years (Faceted)

```{r neighbourhood-facet, fig.height=10, fig.width=12}
neighbourhood_data <- crime_data %>%
  filter(YEAR %in% c(2018, 2019, 2023),
         NEIGHBOURHOOD != "Central Business District") %>%
  count(YEAR, NEIGHBOURHOOD)

ggplot(neighbourhood_data,
       aes(x = reorder(NEIGHBOURHOOD, n), y = n, fill = n)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ YEAR, scales = "free_x", ncol = 3) +
  scale_fill_viridis_c(option = "D") +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Total Crimes by Neighbourhood",
    subtitle = "Side-by-side comparison: 2018, 2019, and 2023 (CBD excluded)",
    x = NULL, y = "Total Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(axis.text.y = element_text(size = 7.5))
```

## Crime Concentration Across Neighbourhoods

How concentrated is crime? The chart below shows cumulative crime share as we move
from the lowest-crime to the highest-crime neighbourhood. A steep curve at the right
indicates that a small number of neighbourhoods account for a large share of crime.

```{r crime-concentration, fig.height=6}
hood_concentration <- crime_data %>%
  filter(YEAR == 2023) %>%
  count(NEIGHBOURHOOD, sort = TRUE) %>%
  arrange(n) %>%
  mutate(
    cumulative_n = cumsum(n),
    cumulative_pct = cumulative_n / sum(n) * 100,
    hood_pct = row_number() / n() * 100
  )

# Find the 80/20 point
pct_80 <- hood_concentration %>%
  filter(cumulative_pct >= 80) %>%
  slice(1)
hoods_for_80 <- round(100 - pct_80$hood_pct, 0)

ggplot(hood_concentration, aes(x = hood_pct, y = cumulative_pct)) +
  geom_line(color = bc_accent, linewidth = 1.3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_hline(yintercept = 80, linetype = "dotted", color = bc_red, alpha = 0.6) +
  annotate("text", x = 20, y = 83,
           label = glue::glue("80% of crime from ~{hoods_for_80}% of neighbourhoods"),
           color = bc_red, size = 3.5, hjust = 0) +
  scale_x_continuous(labels = function(x) paste0(x, "%")) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Crime Concentration Curve (2023)",
    subtitle = "Cumulative % of crime vs cumulative % of neighbourhoods (Lorenz-style)",
    x = "Cumulative % of Neighbourhoods (low to high crime)",
    y = "Cumulative % of Total Crime",
    caption = "Dashed diagonal = perfect equality. Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(panel.grid.major.x = element_line(color = "grey92"))
```

## Neighbourhood Crime Change: 2018 vs 2023

Which neighbourhoods saw the largest increases or decreases in crime between 2018
and 2023? The diverging bar chart below makes this immediately clear.

```{r neighbourhood-change, fig.height=8}
hood_change <- crime_data %>%
  filter(YEAR %in% c(2018, 2023),
         NEIGHBOURHOOD != "Central Business District") %>%
  count(YEAR, NEIGHBOURHOOD) %>%
  pivot_wider(names_from = YEAR, values_from = n, values_fill = 0) %>%
  mutate(
    change_pct = safe_pct_change(`2023`, `2018`),
    abs_change = `2023` - `2018`,
    direction  = ifelse(change_pct > 0, "Increase", "Decrease")
  ) %>%
  filter(!is.na(change_pct)) %>%
  arrange(change_pct)

ggplot(hood_change,
       aes(x = reorder(NEIGHBOURHOOD, change_pct), y = change_pct, fill = direction)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.6, color = "grey40") +
  geom_text(
    aes(label = paste0(ifelse(change_pct > 0, "+", ""), change_pct, "%")),
    hjust = ifelse(hood_change$change_pct > 0, -0.1, 1.1),
    size = 3, fontface = "bold"
  ) +
  coord_flip() +
  scale_fill_manual(values = c("Increase" = bc_red, "Decrease" = bc_green)) +
  scale_y_continuous(labels = function(x) paste0(x, "%"),
                     expand = expansion(mult = 0.15)) +
  labs(
    title = "Crime Change by Neighbourhood: 2018 vs 2023",
    subtitle = "Percentage change in total reported crimes; red = increase, green = decrease",
    x = NULL, y = "Change (%)",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r neighbourhood-change-table}
kable(
  hood_change %>%
    select(NEIGHBOURHOOD, `2018`, `2023`, abs_change, change_pct, direction) %>%
    rename(
      Neighbourhood = NEIGHBOURHOOD,
      `Abs. Change` = abs_change,
      `Change (%)` = change_pct,
      Trend = direction
    ) %>%
    arrange(desc(`Change (%)`)),
  caption = "Table 6: Crime volume change by neighbourhood, 2018 to 2023",
  format.args = list(big.mark = ",")
)
```

## Neighbourhood Crime Profiles (2023)

Different neighbourhoods may have different *types* of crime, not just different
volumes. This heatmap shows what fraction of each neighbourhood's crime falls into
each category, revealing distinct neighbourhood "profiles."

```{r hood-profiles, fig.height=10, fig.width=12}
# Use top 5 neighbourhoods and top 6 crime types for readability
top5_hoods <- crime_data %>%
  filter(YEAR == 2023, NEIGHBOURHOOD != "Central Business District") %>%
  count(NEIGHBOURHOOD, sort = TRUE) %>%
  head(8) %>%
  pull(NEIGHBOURHOOD)

top6_types <- crime_data %>%
  filter(YEAR == 2023) %>%
  count(TYPE, sort = TRUE) %>%
  head(6) %>%
  pull(TYPE)

hood_profile <- crime_data %>%
  filter(YEAR == 2023, NEIGHBOURHOOD %in% top5_hoods, TYPE %in% top6_types) %>%
  count(NEIGHBOURHOOD, TYPE) %>%
  group_by(NEIGHBOURHOOD) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  ungroup()

ggplot(hood_profile, aes(x = TYPE, y = NEIGHBOURHOOD, fill = pct)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(aes(label = paste0(pct, "%"),
                color = pct > median(pct)),
            size = 3.2, fontface = "bold", show.legend = FALSE) +
  scale_color_manual(values = c("TRUE" = "white", "FALSE" = "grey20")) +
  scale_fill_viridis_c(option = "D", name = "Share (%)") +
  labs(
    title = "Neighbourhood Crime Profiles (2023)",
    subtitle = "What % of each neighbourhood's crime falls into each type?",
    x = NULL, y = NULL,
    caption = "Top 8 neighbourhoods x top 6 crime types. Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(
    panel.grid = element_blank(),
    legend.position = "right",
    axis.text.x = element_text(angle = 35, hjust = 1, size = 9)
  )
```

---

# Part 4: Seasonal and Monthly Patterns

Crime is not uniformly distributed throughout the year. Understanding monthly
patterns can inform resource allocation and public safety planning.

```{r monthly-check, include=FALSE}
has_month <- "MONTH" %in% names(crime_data)
```

```{r monthly-analysis, eval=has_month, fig.height=7}
month_labels <- c("Jan","Feb","Mar","Apr","May","Jun",
                   "Jul","Aug","Sep","Oct","Nov","Dec")

monthly_crime <- crime_data %>%
  count(YEAR, MONTH) %>%
  mutate(MONTH_label = factor(month_labels[MONTH], levels = month_labels))

monthly_avg <- monthly_crime %>%
  group_by(MONTH, MONTH_label) %>%
  summarise(avg_n = mean(n), .groups = "drop")

ggplot(monthly_crime, aes(x = MONTH_label, y = n, group = YEAR, color = factor(YEAR))) +
  geom_line(alpha = 0.5, linewidth = 0.7) +
  geom_line(data = monthly_avg, aes(x = MONTH_label, y = avg_n, group = 1),
            color = "black", linewidth = 1.5, linetype = "solid", inherit.aes = FALSE) +
  scale_color_brewer(palette = "Set1", name = "Year") +
  labs(
    title = "Monthly Crime Patterns by Year",
    subtitle = "Thin coloured lines = individual years; thick black line = multi-year average",
    x = NULL, y = "Reported Crimes",
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(legend.position = "right")
```

```{r monthly-heatmap, eval=has_month, fig.height=6}
ggplot(monthly_crime, aes(x = MONTH_label, y = factor(YEAR), fill = n)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = comma(n)), size = 3, color = "white", fontface = "bold") +
  scale_fill_viridis_c(option = "B", name = "Crimes", labels = comma) +
  labs(
    title = "Crime Volume Heatmap: Month x Year",
    subtitle = "Darker cells indicate higher crime volume",
    x = NULL, y = NULL,
    caption = "Source: BC municipal crime data"
  ) +
  theme_bc_crime() +
  theme(panel.grid = element_blank(), legend.position = "right")
```

```{r no-month-note, eval=!has_month, results='asis'}
cat("*Monthly analysis is not available because the dataset does not contain a MONTH column.*\n\n")
```

---

# Part 5: Crime Severity Index

Not all crimes carry equal weight. A simple severity-weighted index gives a more
nuanced view than raw counts alone. The weights below are illustrative and based
on common criminal justice severity scales.

```{r severity-weights-table}
severity_weights <- data.frame(
  TYPE = c("Homicide", "Offence Against a Person",
           "Break and Enter Commercial", "Break and Enter Residential/Other",
           "Vehicle Collision or Pedestrian Struck (with Fatality)",
           "Vehicle Collision or Pedestrian Struck (with Injury)"),
  weight = c(10, 7, 5, 5, 9, 4),
  stringsAsFactors = FALSE
)

kable(
  severity_weights %>% arrange(desc(weight)) %>%
    rename(`Crime Type` = TYPE, `Severity Weight` = weight),
  caption = "Table 8: Severity weight assignments (all other types default to 2)"
)
```

```{r severity-index, fig.height=6}

crime_with_weight <- crime_data %>%
  left_join(severity_weights, by = "TYPE") %>%
  mutate(weight = ifelse(is.na(weight), 2, weight))

severity_by_year <- crime_with_weight %>%
  group_by(YEAR) %>%
  summarise(
    total_crimes = n(),
    severity_score = sum(weight),
    avg_severity = round(mean(weight), 2),
    .groups = "drop"
  )

ggplot(severity_by_year, aes(x = YEAR)) +
  geom_line(aes(y = avg_severity), color = bc_red, linewidth = 1.3) +
  geom_point(aes(y = avg_severity), color = bc_red, size = 3.5,
             shape = 21, fill = "white", stroke = 1.5) +
  geom_text(aes(y = avg_severity, label = avg_severity),
            vjust = -1.3, size = 3.5, color = "grey30") +
  scale_x_continuous(breaks = severity_by_year$YEAR) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  labs(
    title = "Average Crime Severity Index by Year",
    subtitle = "Weighted index: Homicide=10, Person offences=7, B&E=5, default=2",
    x = NULL, y = "Average Severity Score",
    caption = "Note: Weights are illustrative. Source: BC municipal crime data"
  ) +
  theme_bc_crime()
```

```{r severity-table}
kable(
  severity_by_year %>%
    rename(Year = YEAR, `Total Crimes` = total_crimes,
           `Severity Sum` = severity_score, `Avg Severity` = avg_severity),
  caption = "Table 7: Annual severity index",
  format.args = list(big.mark = ",")
)
```

---

# Summary and Key Findings

```{r summary-stats, results='asis'}
total_latest  <- crime_data %>% filter(YEAR == max(YEAR)) %>% nrow()
total_earliest <- crime_data %>% filter(YEAR == min(YEAR)) %>% nrow()
overall_change <- safe_pct_change(total_latest, total_earliest)
latest_year    <- max(crime_data$YEAR)
earliest_year  <- min(crime_data$YEAR)

top_type <- crime_data %>%
  filter(YEAR == latest_year) %>%
  count(TYPE, sort = TRUE) %>%
  slice(1)

worst_hood <- hood_change %>% arrange(desc(change_pct)) %>% slice(1)
best_hood  <- hood_change %>% arrange(change_pct) %>% slice(1)

cat(glue::glue("
## Key Takeaways

1. **Overall trajectory:** Between {earliest_year} and {latest_year}, total reported
   crimes changed by **{ifelse(overall_change > 0, '+', '')}{overall_change}%**
   (from {comma(total_earliest)} to {comma(total_latest)} incidents).

2. **Dominant crime type ({latest_year}):** *{top_type$TYPE}* led with
   **{comma(top_type$n)}** incidents.

3. **Biggest neighbourhood increase:** *{worst_hood$NEIGHBOURHOOD}* saw a
   **+{worst_hood$change_pct}%** increase from {earliest_year} to {latest_year}.

4. **Biggest neighbourhood decrease:** *{best_hood$NEIGHBOURHOOD}* saw a
   **{best_hood$change_pct}%** decrease over the same period.

5. **Post-COVID context:** The 2020\u20132023 window captures pandemic-era disruptions
   and the subsequent recovery in reported crime volumes.

6. **Severity nuance:** Raw crime counts do not capture severity. The weighted
   severity index provides a complementary lens showing whether the *mix* of
   offences is becoming more or less serious over time.
"))
```

---

## Methodology Notes

- **Data source:** BC municipal crime statistics (publicly available CSV).
- **Exclusions:** The Central Business District is excluded from neighbourhood
  comparisons to prevent its high volume from compressing the scale.
- **Severity weights:** The crime severity index uses illustrative weights and
  should be calibrated against official severity scales for policy use.
- **Limitations:** This analysis uses reported crime counts, which may not reflect
  actual crime rates due to under-reporting. Population normalization is not applied
  as neighbourhood-level population data was not available in the dataset.

---

*Report generated on `r Sys.Date()` using R `r getRversion()`.*
