---
title: "BC Crime Data - Exploration Worksheet"
author: "Casper Kacper Ruta"
date: "2024-02-25"
output:
  html_document:
    toc: true
    theme: flatly
    code_folding: show
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 9,
  fig.height = 5
)
```

# Purpose

This worksheet is a companion to the main analysis report (`BC CRIME.Rmd`). It
provides a quick-start template for ad-hoc exploration of the BC crime dataset.
Use the code blocks below as building blocks for your own analyses.

# Setup

```{r load-packages}
packages <- c("here", "readr", "stringr", "clock", "dplyr", "ggplot2",
              "rmarkdown", "knitr", "magrittr", "glue", "xfun", "fs",
              "tidyr", "scales")

installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(knitr)
```

```{r load-data}
crime_data <- read.csv("bc_crime.csv")
```

# Quick Summary

```{r summary-overview}
cat("Records:", comma(nrow(crime_data)), "\n")
cat("Years:", paste(range(crime_data$YEAR), collapse = " to "), "\n")
cat("Crime types:", length(unique(crime_data$TYPE)), "\n")
cat("Neighbourhoods:", length(unique(crime_data$NEIGHBOURHOOD)), "\n")
```

```{r structure}
str(crime_data)
```

```{r preview}
kable(head(crime_data, 10), caption = "First 10 records")
```

# Exploration Templates

## Template 1: Crime counts by year

```{r yearly-counts}
crime_data %>%
  count(YEAR) %>%
  ggplot(aes(x = YEAR, y = n)) +
  geom_col(fill = "#08519c", width = 0.6) +
  geom_text(aes(label = comma(n)), vjust = -0.5, size = 3.5) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.12))) +
  scale_x_continuous(breaks = unique(crime_data$YEAR)) +
  labs(title = "Crime Count per Year", x = "Year", y = "Count") +
  theme_minimal()
```

## Template 2: Filter by crime type

```{r homicide-filter}
crime_data %>%
  filter(TYPE == "Homicide") %>%
  count(YEAR) %>%
  ggplot(aes(x = YEAR, y = n)) +
  geom_col(fill = "#c0392b", width = 0.6) +
  geom_text(aes(label = n), vjust = -0.5, fontface = "bold") +
  scale_x_continuous(breaks = unique(crime_data$YEAR)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Homicide Count per Year", x = "Year", y = "Homicides") +
  theme_minimal()
```

## Template 3: Neighbourhood comparison (single year)

```{r neighbourhood-template}
target_year <- 2023

crime_data %>%
  filter(YEAR == target_year, NEIGHBOURHOOD != "Central Business District") %>%
  count(NEIGHBOURHOOD, sort = TRUE) %>%
  head(10) %>%
  ggplot(aes(x = reorder(NEIGHBOURHOOD, n), y = n, fill = n)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_viridis_c(option = "D", direction = -1) +
  scale_y_continuous(labels = comma) +
  labs(
    title = paste("Top 10 Neighbourhoods by Crime Volume (", target_year, ")"),
    x = NULL, y = "Total Crimes"
  ) +
  theme_minimal()
```

## Template 4: Cross-tabulation

```{r cross-tab}
crime_data %>%
  filter(YEAR == 2023) %>%
  count(TYPE, sort = TRUE) %>%
  mutate(share = round(n / sum(n) * 100, 1)) %>%
  kable(caption = "Crime type distribution (2023)",
        format.args = list(big.mark = ","))
```

---

*Use this worksheet as a starting point. Copy and modify the templates above
for your own investigations.*
